### 2021-04-26 submission/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\vspace{35mm}

\hypertarget{an-osmotic-laxative-renders-mice-susceptible-to-prolonged-clostridioides-difficile-colonization-and-hinders-clearance}{%
\section{\texorpdfstring{An osmotic laxative renders mice susceptible to
prolonged \emph{Clostridioides difficile} colonization and hinders
clearance}{An osmotic laxative renders mice susceptible to prolonged Clostridioides difficile colonization and hinders clearance}}\label{an-osmotic-laxative-renders-mice-susceptible-to-prolonged-clostridioides-difficile-colonization-and-hinders-clearance}}

\vspace{35mm}

Sarah Tomkovich\textsuperscript{1}, Ana Taylor\textsuperscript{1}, Jacob
King\textsuperscript{1}, Joanna Colovas\textsuperscript{1}, Lucas
Bishop\textsuperscript{1}, Kathryn McBride\textsuperscript{1}, Sonya
Royzenblat\textsuperscript{1}, Nicholas A. Lesniak\textsuperscript{1},
Ingrid L. Bergin\textsuperscript{2}, Patrick D.
Schloss\textsuperscript{1\(\dagger\)}

\vspace{40mm}

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:pschloss@umich.edu}{\nolinkurl{pschloss@umich.edu}}

1. Department of Microbiology and Immunology, University of Michigan,
Ann Arbor, MI, USA

2. The Unit for Laboratory Animal Medicine, University of Michigan, Ann
Arbor, MI, USA

\newpage
\linenumbers

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

(Modify depending on target journal, currently abstract submitted to
World Microbe Forum) Antibiotics are a major risk factor for
\emph{Clostridioides difficile} infections (CDIs) because of their
impact on the intestinal microbiome. However, non-antibiotic medications
such as the ubiquitous osmotic laxative polyethylene glycol (PEG) 3350,
also alter the microbiota, but whether PEG impacts CDI susceptibility
and clearance is unclear. To examine how PEG impacts susceptibility, we
treated C57Bl/6 mice with 5-day and 1-day doses of 15\% PEG in the
drinking water and then challenged the mice with \emph{C. difficile} 630
spores. We used clindamycin-treated mice as a control because they
consistently clear \emph{C. difficile} within 10 days post-infection
(dpi). To examine how PEG treatment impacts clearance, we administered
PEG for 1 day to clindamycin-treated, \emph{C. difficile}-challenged
mice either immediately following challenge or 3 dpi. We collected
longitudinal stool samples to examine \emph{C. difficile} levels in the
stool via anaerobic culture and profiled the microbiota by 16S rRNA
sequencing. PEG treatment alone was sufficient to render mice
susceptible to CDI and 5-day PEG-treated mice remain colonized for up to
30 dpi. Additionally, 5-day PEG treated mice remained susceptible to CDI
10-days post treatment. In contrast, 1-day PEG treated mice were
transiently colonized, clearing \emph{C. difficile} within 7 dpi.
Although 5-day PEG-treated mice exhibited prolonged \emph{C. difficile}
colonization, we saw no difference in histological inflammation between
PEG- and clindamycin-treated mice. Additionally, administering PEG to
mice after \emph{C. difficile} challenge prolonged colonization up to 30
dpi in mice that received PEG immediately after challenge and 15 dpi in
mice that received PEG 3 dpi. When we examined microbiota composition
across our different treatment groups, we found decreased richness in
the PEG-treated mice that exhibited prolonged \emph{C. difficile}
colonization. Importantly, there were increased Bacteroides and
Enterobacteriaceae and decreased Lachnospiraceae and Oscillibacter in
most of the PEG-treated mice with prolonged \emph{C. difficile}
colonization. Our findings suggest the osmotic laxative PEG 3350 alters
the mouse microbiota and disrupts colonization resistance to \emph{C.
difficile}, as well as clearance in mice with a CDI. Considering that
most hospitals recommend not performing \emph{C. difficile} testing on
patients taking laxatives and laxatives are used when administering
fecal microbiota transplants via colonoscopy to patients with recurrent
CDIs, further studies are needed to evaluate if laxatives impact human
microbiota colonization resistance.

\newpage

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Antibiotics are a major risk factor for \emph{Clostridioides difficile}
infections (CDIs) because they disrupt microbiota colonization
resistance (1). However, antibiotics are not the only types of
medications that disrupt the microbiota (2--4). Although, other
medications (proton pump inhibitors, osmotic laxatives, antimotility
agents, and opioids) have been implicated as risk or protective factors
for CDIs through epidemiological studies, whether the association is due
to their impact on the microbiome is still unclear (5--9).

Many of the non-antibiotic medications associated with CDIs are known to
modulate gastrointestinal motility leading to either increased or
decreased colonic transit time, which in turn also strongly impacts
microbiota composition and function (10, 11). Stool consistency often
serves as an approximation of intestinal motility. Our group has shown
that when \emph{C. difficile} negative controls are separated into two
groups based on stool consistency, there are shared microbiota features
such as lower alpha diversity in samples from CDI patients and control
patients with diarrhea compared to control samples that were \emph{C.
difficile} negative with non-diarrheal consistency (12). These results
led to a hypothesis that bacterial communities from patients
experiencing diarrhea are susceptible to developing CDIs.

Osmotic laxatives can lead to diarrhea depending on the administered
dose and temporarily disrupt the human intestinal microbiota (13). The
ubiquitous osmotic laxative, polyethylene glycol (PEG) 3350 is found in
Miralax, Nulytely, and Golytely and is also commonly used as bowel
preparation for colonoscopies. Interestingly, previous studies have
shown that treating mice with PEG alone altered microbiota composition,
reduced acetate and butyrate production, altered the mucus barrier, and
rendered the mice susceptible to \emph{C. difficile} infection (14--17).
The mucus barrier is thought to mediate protection from \emph{C.
difficile} infections by protecting intestinal epithelial cells from the
toxins produced by \emph{C. difficile} (18, 19). However, whether
laxative administration results in more severe CDIs in mice and how long
mice remain colonized with \emph{C. difficile} after challenge is
unclear.

Beyond susceptibility, PEG is also relevant in the context of treating
recurrent CDIs via fecal microbiota transplant (FMT) where a healthy
microbiota is administered to the patient to restore colonization
resistance. For FMTs that are delivered via colonoscopy, patients
typically undergo bowel preparation by taking an osmotic laxative prior
to the procedure. Many of the FMT studies to date rationalize the use of
laxatives (20--22) based on a 1996 case study with 2 pediatric patients
where the authors suggested in the discussion that the laxative may help
flush \emph{C. difficile} spores and toxins from the intestine (23).

In the past, our group has used C57BL6 mice to characterize how
antibiotics including clindamycin disrupt the microbiota and influence
\emph{C. difficile} susceptibility and clearance (24--26). Although, two
groups have now shown PEG treatment alone renders mice susceptible to
\emph{C. difficile} (15, 17), these studies have raised additional
questions regarding the dynamics and severity of infection as well as
the role of laxative treatment in \emph{C. difficile} clearance that
should be addressed to better inform how we think about laxatives in the
context of CDIs. Here, we used our C57BL/6 clindamycin model as a
control group to characterize how long PEG-treated mice remain
susceptible, whether PEG treatment results in more severe CDI and
sustained \emph{C. difficile} colonization, and if PEG treatment
post-CDI can promote \emph{C. difficile} clearance.

\hypertarget{results}{%
\subsection{Results}\label{results}}

\textbf{5-day laxative treatment leads to prolonged \emph{C. difficile}
colonization in mice.} We compared mice treated with the osmotic
laxative PEG 3350 to our standard 10 mg/kg clindamycin treatment, which
temporarily renders the mice susceptible to \emph{C. difficile}, with
mice typically clearing \emph{C. difficile} within 10 days
post-infection (9, 26). All PEG-treated mice were administered a 15\%
PEG solution in the drinking water for 5-days, one group was also
treated with clindamycin, and one group was allowed to recover for 10
days prior to challenge (Fig. 1A). PEG treatment resulted in weight loss
in all 3 groups of PEG-treated mice, with the greatest change in weight
observed on the fifth day of PEG treatment and the mice recovered most
of the weight five days after treatment (Fig. 1B). After either PEG,
clindamycin, or PEG and clindamycin treatment all mice were challenged
with 10\textsuperscript{3} \emph{C. difficile} 630 spores (Fig. 1A). All
treatments rendered mice susceptible to \emph{C. difficile}
colonization. However, PEG-treated mice remained colonized with \emph{C.
difficile} at a high level through thirty days post-infection (Fig. 1C).
In contrast, the clindamycin-treated mice cleared \emph{C. difficile}
within ten days post-infection. Surprisingly, PEG-treated mice were
still susceptible to \emph{C. difficile} infection after 10-days of
recovery from treatment, although \emph{C. difficile} was not detectable
in most of the group in the initial five days post-infection (Fig. 1C).
From 8 days post-infection onward, the median \emph{C. difficile} CFU
stabilized for the 10-day recovery group of PEG-treated mice and
remained high through 30 days post-infection (Fig. 1C). Thus, osmotic
laxative treatment alone was sufficient to render mice susceptible to
prolonged \emph{C. difficile} colonization and PEG-treated mice remained
susceptible up to ten days post-treatment.

\textbf{5-day laxative treatment differentially disrupts the fecal
microbiota compared to clindamycin treatment.} Since osmotic laxatives
and clindamycin have previously been shown to disrupt the murine
microbiota (14--17), we hypothesized the different \emph{C. difficile}
colonization dynamics between mice treated with the osmotic laxative or
clindamycin were due to the two drugs having differential effects on the
microbiota. We profiled the stool microbiota over time by sequencing the
V4 region of the 16S rRNA gene to compare changes across treatment
groups. We found time and treatment group explained half of the observed
variation between fecal communities with most of the remaining variation
explained by interactions between treatment group and other experimental
variables including time, cage effects, and sequencing preparation plate
(PERMANOVA combined R\textsuperscript{2} = 0.95, \emph{P} \textless{}
0.001, Fig. 2A, Data Set S1, Sheet X). Cage effects refer to the
well-documented phenomenon that mice housed in the same cages have
similar microbial communities due to copraphagy (27). We tried to
minimize the impact of cage effects on our experiment by breaking up
cagemates when assigning mice to treatment groups and primarily housing
only two mice per cage. Importantly, although we conducted a total of 5
separate experiments, the experiment number and it's interaction with
treatment group was not one of the variables that significantly
explained the observed variation in fecal communities (Data Set S1,
Sheet X). Interestingly, none of the treatment groups recovered to their
baseline community structure either 10 or 30 days post-infection
suggesting other community features besides recovery to baseline were
responsible for the prolonged \emph{C. difficile} colonization in
PEG-treated mice (Fig. 2B).

Next, we examined alpha diversity by looking at Shannon diversity over
time. Although both clindamycin and PEG treatments decreased diversity,
Shannon diversity was lower in the groups of mice that received PEG
treatment compared to those that received clindamycin through thirty
days post-infection (Fig. 2C). We next examined the bacterial genera
that shifted after PEG treatment by comparing the baseline samples of
mice treated with only PEG to samples from the same mice one day
post-treatment. We found 18 genera that were altered by PEG treatment
(Data Set S1, Sheet X). The majority of the bacterial relative
abundances decreased after PEG, but \emph{Enterobacteriaceae} and
\emph{Bacteroides} increased and the increase in \emph{Bacteroides} was
unique to PEG treatment, as \emph{Bacteroides} actually decreased in
clindamycin treated mice (Fig. 2D). Finally, we examined the bacteria
that differed across treatment groups over multiple time points. We
found 24 genera were different over multiple time points out of the 33
genera that were different between treatment groups (Fig. 2E, Data Set
S1, Sheet X). Thus, PEG had a significant impact on the fecal microbiota
that was maintained over time and was distinct from clindamycin
treatment.

Interestingly, \emph{C. difficile} was not immediately detectable in the
stools of the PEG-treated mice that were allowed to recover for 10 days
prior to challenge. We decided to examine the bacteria that changed
during the post-infection period when the group median \emph{C.
difficile} shifted from undetectable at 1 day post-infection to
detectable in the stool samples with the median stabilizing around 8
days post-infection (Fig. S1A). Interestingly, we found
\emph{Erysipelotrichaceae}, \emph{Enterobacteriaceae}, and
\emph{Akkermansia} were changing during the time period when \emph{C.
difficile} was becoming detectable in the stools (Fig. S1B), although
none of the bacteria were significant after multiple hypothesis
correction (Data Set S1, Sheet X). While we did not identify a clear
signal to explain the delayed appearance of \emph{C. difficile} in the
5-day PEG mice that were allowed to recover for 10 days prior to
challenge, the delay is striking and could reflect changes in microbial
activity or metabolites that were not examined in this study.

\textbf{5-day laxative treatment does not promote more severe CDIs
despite altering the mucosal microbiota.} Given the findings from a
previous study that demonstrated PEG treatment disrupts the mucus layer
and alters the immune response in mice (16), we decided to examine the
impact of PEG treatment on the mucosal microbiota and CDI severity. To
evaluate the mucosal microbiota, we sequenced snips of tissue collected
from the cecum, proximal colon, and distal colon. Similar to what was
observed with the stool samples, alpha diversity was lower in the
PEG-treated mice compared to clindamycin treatment (Fig. 3A). Although
alpha diversity continued to increase over time based on the communities
from PEG-treated mice collected at 20 and 30 days post-infection (Fig.
3A, Data Set S1, Sheet X). Group, time point, and their interactions
with other variables (cage, experiment number, and sample type)
explained the majority of the variation observed in mucosal communities
(PERMANOVA combined R\textsuperscript{2} = 0.83, \emph{P} \textless{}
0.05, Fig. 3B, Data Set S1, Sheet X). We saw the greatest difference in
the mucosal microbiota between treatment groups at 6 days post-infection
with 16 genera that were significantly different in all three of the
tissue types we collected (cecum, proximal colon, and distal colon; Fig.
S3A, Data Set S1, Sheet X). However, at 30 days post-infection only
\emph{Bacteroides}, \emph{Clostridiales}, \emph{Firmicutes}, and
\emph{Ruminococcaceae} were different between treatment groups and only
in the cecum tissues (Fig. 3C, Fig. 2E, Data Set S1, Sheet X).
Interestingly, one of the bacteria that was different between treatment
groups at 6 days post-infection was \emph{Peptostreptococcaceae} (the
genera with an OTU sequence that matches \emph{C. difficile}).
\emph{Peptostreptococcaceae} was primarily only present in the 5-day PEG
treatment group of mice and decreased in the proximal and distal colon
tissues over time (Fig. S2B). Thus, PEG treatment had a significant
impact on the mucosal microbiota and we detected \emph{C. difficile}
sequences in the cecum, proximal colon, and distal colon tissue
communities.

Next, we examined CDI severity by evaluating cecum and colon
histopathology (28) and found there was no difference in cecum and colon
scores between clindamycin and PEG-treated mice that were challenged
with \emph{C. difficile} at 4 days post-infection (Fig. 3D), the time
point typically examined in \emph{C. difficile} 630 challenged mice
(29). We also looked at 6 days post-infection because that was when we
started to see a large difference in \emph{C. difficile} colonization
levels between PEG- and clindamycin-treated mice (Fig. 1C). Although,
there was a slight difference in the colon between PEG and
clindamycin-treated mice, there was no difference in the cecum and the
overall score is still relatively low given that the max possible
summary score is twelve (Fig. 3E). Therefore, although PEG treatment had
a disruptive effect on the mucosal microbiota, the impact of \emph{C.
difficile} 630 infection on the cecum and colon was similar between PEG
and clindamycin treated mice.

\textbf{\emph{C. difficile} challenge does not have a synergistic
disruptive effect on the microbiota of PEG-treated mice} Because
\emph{C. difficile} itself can have an impact on the microbiota (30), we
also sequenced the tissue and stools of mock-challenged clindamycin and
5-day PEG treated mice. Examining the stools of the mock-challenged mice
revealed similar bacterial disruptions as the \emph{C. difficile}
challenged mice (Fig. S2A-C). Similarly, there was no difference between
the tissues of mock and \emph{C. difficile} challenged mice (Fig.
S2D-F). Thus, most of the microbiota alterations we observed in the
PEG-treated mice were a result of the laxative and not an interaction
between the laxative and \emph{C. difficile}.

\textbf{1-day laxative treatment results in transient \emph{C.
difficile} colonization and minor microbiota disruption} Next, we
decided to examine how a shorter osmotic laxative perturbation would
impact the microbiome and susceptibility to \emph{C. difficile}. We
administered either a 1-day PEG treatment, a 1-day PEG treatment with a
1-day recovery period, or clindamycin to mice before challenging them
with \emph{C. difficile} (Fig. 3A). In contrast to the 5-day PEG treated
mice, the 1-day PEG treated mice were only transiently colonized and
cleared \emph{C. difficile} by 7 days post-infection (Fig. 3B). The
stool communities of PEG-treated mice were also only transiently
disrupted, with Shannon diversity recovering close to baseline by 7 days
post-infection (Fig. 3C-D). We found 14 bacteria were impacted by
treatment, but recovered close to baseline levels by 7 days
post-infection including \emph{Enterobacteriaceae},
\emph{Clostridiales}, \emph{Porpyromonadaceae}, and
\emph{Ruminococcaceae} (Fig. 3E, Data Set S1, Sheet X). These findings
suggest the duration of treatment matters when considering the impact of
the laxatives, with shorter treatments resulting in a transient loss of
\emph{C. difficile} colonization resistance.

\textbf{Post-CDI laxative treatment disrupts clearance in
clindamycin-treated mice regardless of whether an FMT is also
administered} Since 1-day PEG treatment resulted in a more mild
microbiota perturbation, we decided to use the 1-day treatment to
examine the hypothesis that PEG helps to flush \emph{C. difficile}
spores from the intestine. We were also interested in exploring whether
PEG might help with engraftment in the context of FMTs. To examine the
impact of PEG treatment on \emph{C. difficile} clearance and FMT
treatment, we treated 4 groups of mice with clindamycin and then
challenged all mice with \emph{C. difficile} before administering the
following treatments: no additional treatment, 1-day PEG immediately
after challenge, and 1-day PEG treatment 3 days after challenge followed
by either administration of an FMT or PBS solution by oral gavage (Fig.
5A). Contrary to our hypothesis, all groups of mice that received PEG
exhibited prolonged \emph{C. difficile} colonization (Fig. 5B). Next we
examined how post-CDI PEG treatment impacted the microbiota. Alpha
diversity was lower in the PEG-treated mice with the exception of the
PEG-treated mice that were administered an FMT (Fig 5C-D). The FMT
appeared to partially restore Shannon diversity but not richness (Fig.
5C-D). Similarly, we saw some overlap between the communities of mice
that received FMT and the mice treated with only clindamycin after 5
days post-infection (Fig. 6A). The increase in Shannon diversity
suggests that the FMT did have an impact on the microbiota, despite
seeing prolonged \emph{C. difficile} colonization in the FMT treated
mice. However, only \emph{Bacteroidales} and \emph{Porphyromonadaceae}
consistently differed between the mice received either an FMT or PBS
gavage (Fig. 6B), suggesting the FMT only restored a couple of genera.
Overall, we found there were 24 bacteria that differed over multiple
time points between treatment groups including increased
\emph{Akkermansia} and decreased \emph{Ruminococcaceae},
\emph{Clostridiales}, \emph{Lachnospiraceae}, and \emph{Oscillibacter}
in mice that received PEG after \emph{C. difficile} challenge (Fig. 6C).
In sum, administering PEG actually prolonged \emph{C. difficile}
colonization, including in mice that received an FMT, which only
restored a couple of bacterial genera.

\textbf{Five-day post-infection community data can predict which mice
that will have prolonged \emph{C. difficile} colonization} After
identifying bacteria associated with the 5-day, 1-day and post-CDI 1-day
PEG treatments, we decided to examine the bacteria that influenced
prolonged \emph{C. difficile} colonization. We trained 3 types of
machine learning models (random forest, logistic regression, and support
vector machine) with input bacterial community data from 5 days
post-infection to predict whether the mice were still colonized with
\emph{C. difficile} 10 days post-infection. We chose 5 days
post-infection because that was the earliest time point where we would
see a treatment effect in the mice that were given 1-day PEG treatment
three days after \emph{C. difficile} challenge and then administered an
FMT or PBS gavage. The random forest model had the highest performance
(AUROC = 0.90, Data Set S1, Sheet X), so we next performed permutation
importance to examine the bacteria that were the top contributors to the
random forest model predicting prolonged \emph{C. difficile
}colonization. We selected the top 10 bacteria contributing to our
models performance (Fig. 7A) and examined their relative abundance at 5
days post-infection, the time point used to predict \emph{C. difficile}
colonization status on day 10 (Fig. 7B). Next, we focused on the 5
genera that had a greater than 1 \% relative abundance in either the
cleared or colonized mice and examined how the bacteria changed over
time. We found \emph{Enterobacteriaceae} and \emph{Bacteroides} tended
to have a higher relative abundance, \emph{Akkermansia} was initially
decreased and then increased, and \emph{Porphyromonadaceae} and
\emph{Lachnospiraceae} had a lower relative abundance in the mice with
prolonged colonization compared to the mice that cleared \emph{C.
difficile} (Fig. 7C). Together these results suggest a combination of
low and high abundance bacterial genera influence the prolonged
colonization observed in 5-day PEG and post-CDI PEG treated mice.

Previous work examining the impact of PEG on the murine microbiota found
that PEG treatment resulted in the permanent loss of S24-7, also known
as \emph{Muribaculum intestinale} (16). We decided to check our
\emph{Porphyromonadaceae} OTUs because \emph{Muribaculum intestinale} is
classified as \emph{Porphyromonadaceae} by the Ribosomal Database
Project (RDP) database (31) and \emph{Porphyromonadaceae} was a top
feature in the random forest model predicting prolonged \emph{C.
difficile} colonization. We identified 4 OTUs that had at least 92\%
identity to \emph{Muribaculum intestinale} and examined their abundance
in mice that either cleared or were still colonized with \emph{C.
difficile} at 10 days post-infection. While all of the OTUs, were
decreased by PEG and clindamycin treatment, there was some recovery in
the mice that cleared (Fig. S4A). We also examined other
\emph{Porphyromonadaceae} and \emph{Lachnospiraceae} OTUs since these
were the 2 genera that were important to our classification model and
contained multiple OTUs that were different at 5 days post-infection
between mice that either cleared or remained colonized with \emph{C.
difficile} by 10 days post-infection (Data Set S1, sheet X). While
individual \emph{Porphyromonadaceae} and \emph{Lachnospiraceae} OTUs
tended to be more abundant in the mice that clear \emph{C. difficile}
relative to the mice that exhibit prolonged colonization (Fig. S4B-C),
there is no single OTU that fits the pattern we observed at the genus
level (Fig. 7C), suggesting multiple \emph{Porphyromonadaceae} and
\emph{Lachnospiraceae} OTUs influenced \emph{C. difficile} clearance.
Overall, our results suggest that specific bacterial community
differences explain the prolonged \emph{C. difficile} colonization we
observed in 5-day PEG and post-CDI 1-day PEG treated mice.

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

\begin{itemize}
\item
  Summary of major findings (Fig. 8A)
\item
  Discussion of prolonged persistence. \emph{C. difficile} sequences
  detected in tissue samples. Association with mucin-degrading bacteria
  suggested by recent papers.
\item
  Discuss why we might not have observed more severe histology in PEG
  mice relative to clindamycin-treated mice

  \begin{itemize}
  \tightlist
  \item
    Antibiotics may also impact mucus layer
  \item
    Strain of bacteria used
  \end{itemize}
\item
  Protective bacteria missing in PEG-treated mice
\item
  Discuss what these findings might mean for human patients (Fig. 8B)

  \begin{itemize}
  \tightlist
  \item
    What's known regarding laxatives and susceptibility to CDIs
  \item
    Clinical trial of PEG, results never posted (32)
  \item
    Relevance to human FMTs? Unclear what the best administration route
    is because there have been no studies designed to evaluate the best
    administration route for FMTs.
  \end{itemize}
\end{itemize}

\hypertarget{conclusions}{%
\subsection{Conclusions}\label{conclusions}}

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

We thank members of the Schloss lab for feedback on planning the
experiments and data presentation. We thank Andrew Henry for help with
media preparation and bacterial culture and the Microbiology and
Immunology department's postdoc association writing group members for
their feedback on manuscript drafts. We also thank the Unit for
Laboratory Animal Medicine at the University of Michigan for maintaining
our mouse colony and providing the institutional support for our mouse
experiments. Finally, we thank Kwi Kim, Austin Campbell, and Kimberly
Vendrov for their help in maintaining the Schloss lab's anaerobic
chamber. This work was supported by the National Institutes of Health
(U01AI124255). ST was supported by the Michigan Institute for Clincial
and Health Research Postdoctoral Translation Scholars Program
(UL1TR002240 from the National Center for Advancing Translational
Sciences).

\hypertarget{materials-and-methods}{%
\subsection{Materials and Methods}\label{materials-and-methods}}

\begin{itemize}
\tightlist
\item
  Histopathology (33)
\end{itemize}

\newpage

\includegraphics{figure_1.pdf}

\textbf{Figure 1. 5-day PEG treatment prolongs susceptibility and mice
become persistently colonized with \emph{C. difficile}.} A. Setup of the
experimental timeline for experiments with 5-day PEG treated mice.
Clindamycin was administered at 10 mg/kg by intraperitoneal injection.
15\% PEG 3350 was administered in the drinking water for five days. B.
Weight change from baseline weight in groups after treatment with PEG
and/or clindamycin, followed by \emph{C. difficile} challenge. C.
\emph{C. difficile} CFU/gram stool measured over time (N = 16-59 mice
per time point) via serial dilutions. The black line represents the
limit of detection for the first serial dilution. CFU quantification
data was not available for each mouse due to stool sampling difficulties
(particularly the day the mice came off of the PEG treatment) or early
deaths. For B-C, lines represent the median for each treatment group and
circles represent samples from individual mice. Asterisks indicate time
points where the weight change or CFU/g was significantly different
between groups by the Kruskal-Wallis test with Benjamini-Hochberg
correction for testing multiple time points. The data presented are from
a total of 5 separate experiments. \newpage

\includegraphics{figure_2.pdf} \textbf{Figure 2. 5-day PEG treatment
disrupts the stool microbiota for a longer amount of time compared to
clindamycin-treated mice.} A. PCoA of Bray-Curtis distances from stool
samples collected throughout the experiment. \newpage

\includegraphics{figure_3.pdf} \textbf{Figure 3. 5-day PEG treatment
does not result in more severe CDIs, although mucosal microbiota is
altered.} A. \newpage

\includegraphics{figure_4.pdf} \textbf{Figure 4. 1-day PEG treatment
renders mice susceptible to transient \emph{C. difficile} colonization.}
A. Setup of the experimental timeline for the 1-day PEG treated subset
of mice. B. CFU/gram stool measured over time (N = 12-18 mice per time
point) via several dilutions. The black dotted line represents the limit
of detection for the first serial dilution. Asterisks indicate time
points where the CFU/gram was significantly different between groups
using the Kruskall-Wallis test with a Benjamini-Hochberg correction for
multiple time points. C. Principle Coordinate Analysis plot of the
groups over time with the alpha representating the same time scale as in
panel D (day: R\textsuperscript{2} = 0.43; group: R\textsuperscript{2} =
0.19). D. Shannon diversity Index of the groups over time. Only days
with samples from all groups are shown. E. Line plots of relative
percent abundance of selected genera over time. Only days with samples
from all groups shown. The gray line represents the limit of detection.

\includegraphics{figure_5.pdf} \textbf{Figure 5. 1-day PEG treatment
post \emph{C. difficile} challenge prolongs colonization regardless of
whether an FMT is also administered.} All mice were administered 10
mg/kg clindamycin intraperitoneally (IP) 1 day before challenge with
\emph{C. difficile} 630 spores on day 0. 15\% PEG 3350 was administered
in the drinking water for 1 day. Mice feces were collected daily through
the end of the experiment (30 days post-infection for Clind. + 1-day PEG
group, 15 days post-infection for all others). CFU quantification data
was not available for each mouse due to stool sampling difficulties
(particularly the day the mice came off of the PEG treatment) or early
deaths. B. CFU/g of \emph{C. difficile} stool measured over time via
serial dilutions. C. Median and individual Shannon Diversity measured
over time. D. Median and individual richness measured over time. B-D.
Asterisks indicate time points with significant differences between
groups using Kruskall-Wallis tests with a Benjamini-Hochberg correction
for testing multiple times points. Black line represents the limit of
detection for the first serial dilution. Background shading indicates
PEG treatment period for applicable groups. Opacity of points reflects
time point. The data presented are from a total of *** separate
experiments. \newpage

\includegraphics{figure_6.pdf} \textbf{Figure 6. For 1-day PEG treatment
post \emph{C. difficile} challenge mice that also receive an FMT only
some bacterial genera were restored.} A. PCoA of all groups + FMT
plotted over time. Opacity of points reflects time point. B. Relative
abundance of genera significantly different over multiple time points
post FMT or PBS treatment, plotted over time. Limit of detection is
.1\%. C. Relative abundances of top 6 significant genera ranked by
number of days significant, plotted over time. Limit of detection is
.1\%.

\newpage

\includegraphics{figure_7.pdf} \textbf{Figure 7. Specific microbiota
features associated with prolonged \emph{C. difficile} colonization in
PEG treated mice.} A. \newpage

\includegraphics{figure_8.pdf} \textbf{Figure 8. Schematic summarizing
findings.} A. \newpage

\includegraphics{figure_S1.pdf} \textbf{Figure S1. 5-day PEG treatment
plus 10-day recovery mice microbiota dynamics post-infection.} A.
\newpage

\includegraphics{figure_S2.pdf} \textbf{Figure S2. Impact of PEG
treatment on the mucosal microbiota is greatest 6 days post-infection}\\
\newpage

\includegraphics{figure_S3.pdf} \textbf{Figure S3. \emph{C. difficile}
challenge does not enhance the disruptive effect of PEG on the
microbiota.} A. \newpage

\includegraphics{figure_S4.pdf} \textbf{Figure S4. Specific OTUs
associated with clearance that are mostly absent in mice with prolonged
\emph{C. difficile} colonization. Ex. \emph{Muribaculum intestinale}.}
A. \newpage

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Britton2014}{}%
1. \textbf{Britton RA}, \textbf{Young VB}. 2014. Role of the intestinal
microbiota in resistance to colonization by clostridium difficile.
Gastroenterology \textbf{146}:1547--1553.
doi:\href{https://doi.org/10.1053/j.gastro.2014.01.059}{10.1053/j.gastro.2014.01.059}.

\leavevmode\hypertarget{ref-Maier2018}{}%
2. \textbf{Maier L}, \textbf{Pruteanu M}, \textbf{Kuhn M},
\textbf{Zeller G}, \textbf{Telzerow A}, \textbf{Anderson EE},
\textbf{Brochado AR}, \textbf{Fernandez KC}, \textbf{Dose H},
\textbf{Mori H}, \textbf{Patil KR}, \textbf{Bork P}, \textbf{Typas A}.
2018. Extensive impact of non-antibiotic drugs on human gut bacteria.
Nature \textbf{555}:623--628.
doi:\href{https://doi.org/10.1038/nature25979}{10.1038/nature25979}.

\leavevmode\hypertarget{ref-LeBastard2017}{}%
3. \textbf{Bastard QL}, \textbf{Al-Ghalith GA}, \textbf{Grégoire M},
\textbf{Chapelet G}, \textbf{Javaudin F}, \textbf{Dailly E},
\textbf{Batard E}, \textbf{Knights D}, \textbf{Montassier E}. 2017.
Systematic review: Human gut dysbiosis induced by non-antibiotic
prescription medications. Alimentary Pharmacology \& Therapeutics
\textbf{47}:332--345.
doi:\href{https://doi.org/10.1111/apt.14451}{10.1111/apt.14451}.

\leavevmode\hypertarget{ref-VichVila2020}{}%
4. \textbf{Vila AV}, \textbf{Collij V}, \textbf{Sanna S}, \textbf{Sinha
T}, \textbf{Imhann F}, \textbf{Bourgonje AR}, \textbf{Mujagic Z},
\textbf{Jonkers DMAE}, \textbf{Masclee AAM}, \textbf{Fu J},
\textbf{Kurilshikov A}, \textbf{Wijmenga C}, \textbf{Zhernakova A},
\textbf{Weersma RK}. 2020. Impact of commonly used drugs on the
composition and metabolic function of the gut microbiota. Nature
Communications \textbf{11}.
doi:\href{https://doi.org/10.1038/s41467-019-14177-z}{10.1038/s41467-019-14177-z}.

\leavevmode\hypertarget{ref-Oh2018}{}%
5. \textbf{Oh J}, \textbf{Makar M}, \textbf{Fusco C}, \textbf{McCaffrey
R}, \textbf{Rao K}, \textbf{Ryan EE}, \textbf{Washer L}, \textbf{West
LR}, \textbf{Young VB}, \textbf{Guttag J}, \textbf{Hooper DC},
\textbf{Shenoy ES}, \textbf{Wiens J}. 2018. A generalizable, data-driven
approach to predict daily risk ofClostridium difficileInfection at two
large academic health centers. Infection Control \& Hospital
Epidemiology \textbf{39}:425--433.
doi:\href{https://doi.org/10.1017/ice.2018.16}{10.1017/ice.2018.16}.

\leavevmode\hypertarget{ref-Mora2012}{}%
6. \textbf{Mora AL}, \textbf{Salazar M}, \textbf{Pablo-Caeiro J},
\textbf{Frost CP}, \textbf{Yadav Y}, \textbf{DuPont HL}, \textbf{Garey
KW}. 2012. Moderate to high use of opioid analgesics are associated with
an increased risk of clostridium difficile infection. The American
Journal of the Medical Sciences \textbf{343}:277--280.
doi:\href{https://doi.org/10.1097/maj.0b013e31822f42eb}{10.1097/maj.0b013e31822f42eb}.

\leavevmode\hypertarget{ref-Nehra2018}{}%
7. \textbf{Nehra AK}, \textbf{Alexander JA}, \textbf{Loftus CG},
\textbf{Nehra V}. 2018. Proton pump inhibitors: Review of emerging
concerns. Mayo Clinic Proceedings \textbf{93}:240--246.
doi:\href{https://doi.org/10.1016/j.mayocp.2017.10.022}{10.1016/j.mayocp.2017.10.022}.

\leavevmode\hypertarget{ref-Krishna2013}{}%
8. \textbf{Krishna SG}, \textbf{Zhao W}, \textbf{Apewokin SK},
\textbf{Krishna K}, \textbf{Chepyala P}, \textbf{Anaissie EJ}. 2013.
Risk factors, preemptive therapy, and antiperistaltic agents
forClostridium difficileinfection in cancer patients. Transplant
Infectious Disease n/a--n/a.
doi:\href{https://doi.org/10.1111/tid.12112}{10.1111/tid.12112}.

\leavevmode\hypertarget{ref-Tomkovich2019}{}%
9. \textbf{Tomkovich S}, \textbf{Lesniak NA}, \textbf{Li Y},
\textbf{Bishop L}, \textbf{Fitzgerald MJ}, \textbf{Schloss PD}. 2019.
The proton pump inhibitor omeprazole does not promote
\emph{Clostridioides difficile} colonization in a murine model. mSphere
\textbf{4}.
doi:\href{https://doi.org/10.1128/msphere.00693-19}{10.1128/msphere.00693-19}.

\leavevmode\hypertarget{ref-Vandeputte2015}{}%
10. \textbf{Vandeputte D}, \textbf{Falony G}, \textbf{Vieira-Silva S},
\textbf{Tito RY}, \textbf{Joossens M}, \textbf{Raes J}. 2015. Stool
consistency is strongly associated with gut microbiota richness and
composition, enterotypes and bacterial growth rates. Gut
\textbf{65}:57--62.
doi:\href{https://doi.org/10.1136/gutjnl-2015-309618}{10.1136/gutjnl-2015-309618}.

\leavevmode\hypertarget{ref-VujkovicCvijin2020}{}%
11. \textbf{Vujkovic-Cvijin I}, \textbf{Sklar J}, \textbf{Jiang L},
\textbf{Natarajan L}, \textbf{Knight R}, \textbf{Belkaid Y}. 2020. Host
variables confound gut microbiota studies of human disease. Nature
\textbf{587}:448--454.
doi:\href{https://doi.org/10.1038/s41586-020-2881-9}{10.1038/s41586-020-2881-9}.

\leavevmode\hypertarget{ref-Schubert2014}{}%
12. \textbf{Schubert AM}, \textbf{Rogers MAM}, \textbf{Ring C},
\textbf{Mogle J}, \textbf{Petrosino JP}, \textbf{Young VB},
\textbf{Aronoff DM}, \textbf{Schloss PD}. 2014. Microbiome data
distinguish patients with clostridium difficile infection and non-c.
Difficile-associated diarrhea from healthy controls. mBio \textbf{5}.
doi:\href{https://doi.org/10.1128/mbio.01021-14}{10.1128/mbio.01021-14}.

\leavevmode\hypertarget{ref-Nagata2019}{}%
13. \textbf{Nagata N}, \textbf{Tohya M}, \textbf{Fukuda S}, \textbf{Suda
W}, \textbf{Nishijima S}, \textbf{Takeuchi F}, \textbf{Ohsugi M},
\textbf{Tsujimoto T}, \textbf{Nakamura T}, \textbf{Shimomura A},
\textbf{Yanagisawa N}, \textbf{Hisada Y}, \textbf{Watanabe K},
\textbf{Imbe K}, \textbf{Akiyama J}, \textbf{Mizokami M},
\textbf{Miyoshi-Akiyama T}, \textbf{Uemura N}, \textbf{Hattori M}. 2019.
Effects of bowel preparation on the human gut microbiome and metabolome.
Scientific Reports \textbf{9}.
doi:\href{https://doi.org/10.1038/s41598-019-40182-9}{10.1038/s41598-019-40182-9}.

\leavevmode\hypertarget{ref-Kashyap2013}{}%
14. \textbf{Kashyap PC}, \textbf{Marcobal A}, \textbf{Ursell LK},
\textbf{Larauche M}, \textbf{Duboc H}, \textbf{Earle KA},
\textbf{Sonnenburg ED}, \textbf{Ferreyra JA}, \textbf{Higginbottom SK},
\textbf{Million M}, \textbf{Tache Y}, \textbf{Pasricha PJ},
\textbf{Knight R}, \textbf{Farrugia G}, \textbf{Sonnenburg JL}. 2013.
Complex interactions among diet, gastrointestinal transit, and gut
microbiota in humanized mice. Gastroenterology \textbf{144}:967--977.
doi:\href{https://doi.org/10.1053/j.gastro.2013.01.047}{10.1053/j.gastro.2013.01.047}.

\leavevmode\hypertarget{ref-Ferreyra2014}{}%
15. \textbf{Ferreyra JA}, \textbf{Wu KJ}, \textbf{Hryckowian AJ},
\textbf{Bouley DM}, \textbf{Weimer BC}, \textbf{Sonnenburg JL}. 2014.
Gut microbiota-produced succinate promotes c.~difficile infection after
antibiotic treatment or motility disturbance. Cell Host \& Microbe
\textbf{16}:770--777.
doi:\href{https://doi.org/10.1016/j.chom.2014.11.003}{10.1016/j.chom.2014.11.003}.

\leavevmode\hypertarget{ref-Tropini2018}{}%
16. \textbf{Tropini C}, \textbf{Moss EL}, \textbf{Merrill BD},
\textbf{Ng KM}, \textbf{Higginbottom SK}, \textbf{Casavant EP},
\textbf{Gonzalez CG}, \textbf{Fremin B}, \textbf{Bouley DM},
\textbf{Elias JE}, \textbf{Bhatt AS}, \textbf{Huang KC},
\textbf{Sonnenburg JL}. 2018. Transient osmotic perturbation causes
long-term alteration to the gut microbiota. Cell
\textbf{173}:1742--1754.e17.
doi:\href{https://doi.org/10.1016/j.cell.2018.05.008}{10.1016/j.cell.2018.05.008}.

\leavevmode\hypertarget{ref-VanInsberghe2020}{}%
17. \textbf{VanInsberghe D}, \textbf{Elsherbini JA}, \textbf{Varian B},
\textbf{Poutahidis T}, \textbf{Erdman S}, \textbf{Polz MF}. 2020.
Diarrhoeal events can trigger long-term clostridium difficile
colonization with recurrent blooms. Nature Microbiology
\textbf{5}:642--650.
doi:\href{https://doi.org/10.1038/s41564-020-0668-2}{10.1038/s41564-020-0668-2}.

\leavevmode\hypertarget{ref-Olson2013}{}%
18. \textbf{Olson A}, \textbf{Diebel LN}, \textbf{Liberati DM}. 2013.
Effect of host defenses on clostridium difficile toxininduced intestinal
barrier injury. Journal of Trauma and Acute Care Surgery
\textbf{74}:983--990.
doi:\href{https://doi.org/10.1097/ta.0b013e3182858477}{10.1097/ta.0b013e3182858477}.

\leavevmode\hypertarget{ref-Diebel2014}{}%
19. \textbf{Diebel LN}, \textbf{Liberati DM}. 2014. Reinforcement of the
intestinal mucus layer protects against clostridium difficile intestinal
injury in~vitro. Journal of the American College of Surgeons
\textbf{219}:460--468.
doi:\href{https://doi.org/10.1016/j.jamcollsurg.2014.05.005}{10.1016/j.jamcollsurg.2014.05.005}.

\leavevmode\hypertarget{ref-vanNood2013}{}%
20. \textbf{Nood E van}, \textbf{Vrieze A}, \textbf{Nieuwdorp M},
\textbf{Fuentes S}, \textbf{Zoetendal EG}, \textbf{Vos WM de},
\textbf{Visser CE}, \textbf{Kuijper EJ}, \textbf{Bartelsman JFWM},
\textbf{Tijssen JGP}, \textbf{Speelman P}, \textbf{Dijkgraaf MGW},
\textbf{Keller JJ}. 2013. Duodenal infusion of donor feces for
RecurrentClostridium difficile. New England Journal of Medicine
\textbf{368}:407--415.
doi:\href{https://doi.org/10.1056/nejmoa1205037}{10.1056/nejmoa1205037}.

\leavevmode\hypertarget{ref-Razik2017}{}%
21. \textbf{Razik R}, \textbf{Osman M}, \textbf{Lieberman A},
\textbf{Allegretti JR}, \textbf{Kassam Z}. 2017. Faecal microbiota
transplantation for clostridium difficile infection: A multicentre study
of non-responders. Medical Journal of Australia \textbf{207}:159--160.
doi:\href{https://doi.org/10.5694/mja16.01452}{10.5694/mja16.01452}.

\leavevmode\hypertarget{ref-Postigo2012}{}%
22. \textbf{Postigo R}, \textbf{Kim JH}. 2012. Colonoscopic versus
nasogastric fecal transplantation for the treatment of clostridium
difficile infection: A review and pooled analysis. Infection
\textbf{40}:643--648.
doi:\href{https://doi.org/10.1007/s15010-012-0307-9}{10.1007/s15010-012-0307-9}.

\leavevmode\hypertarget{ref-Liacouras1996}{}%
23. \textbf{Liacouras CA}, \textbf{Piccoli DA}. 1996. Whole-bowel
irrigation as an adjunct to the treatment of chronic, relapsing
clostridium difficile colitis. Journal of Clinical Gastroenterology
\textbf{22}:186--189.
doi:\href{https://doi.org/10.1097/00004836-199604000-00007}{10.1097/00004836-199604000-00007}.

\leavevmode\hypertarget{ref-Schubert2015}{}%
24. \textbf{Schubert AM}, \textbf{Sinani H}, \textbf{Schloss PD}. 2015.
Antibiotic-induced alterations of the murine gut microbiota and
subsequent effects on colonization resistance against \emph{Clostridium
difficile}. mBio \textbf{6}.
doi:\href{https://doi.org/10.1128/mbio.00974-15}{10.1128/mbio.00974-15}.

\leavevmode\hypertarget{ref-Jenior2017}{}%
25. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2017. \emph{Clostridium difficile} colonizes
alternative nutrient niches during infection across distinct murine gut
microbiomes. mSystems \textbf{2}.
doi:\href{https://doi.org/10.1128/msystems.00063-17}{10.1128/msystems.00063-17}.

\leavevmode\hypertarget{ref-Tomkovich2020}{}%
26. \textbf{Tomkovich S}, \textbf{Stough JMA}, \textbf{Bishop L},
\textbf{Schloss PD}. 2020. The initial gut microbiota and response to
antibiotic perturbation influence clostridioides difficile clearance in
mice. mSphere \textbf{5}.
doi:\href{https://doi.org/10.1128/msphere.00869-20}{10.1128/msphere.00869-20}.

\leavevmode\hypertarget{ref-Nguyen2015}{}%
27. \textbf{Nguyen TLA}, \textbf{Vieira-Silva S}, \textbf{Liston A},
\textbf{Raes J}. 2015. How informative is the mouse for human gut
microbiota research? Disease Models \& Mechanisms \textbf{8}:1--16.
doi:\href{https://doi.org/10.1242/dmm.017400}{10.1242/dmm.017400}.

\leavevmode\hypertarget{ref-Reeves2011}{}%
28. \textbf{Reeves AE}, \textbf{Theriot CM}, \textbf{Bergin IL},
\textbf{Huffnagle GB}, \textbf{Schloss PD}, \textbf{Young VB}. 2011. The
interplay between microbiome dynamics and pathogen dynamics in a murine
model of \emph{Clostridium difficile} infection \textbf{2}:145--158.
doi:\href{https://doi.org/10.4161/gmic.2.3.16333}{10.4161/gmic.2.3.16333}.

\leavevmode\hypertarget{ref-Dieterle2020}{}%
29. \textbf{Dieterle MG}, \textbf{Putler R}, \textbf{Perry DA},
\textbf{Menon A}, \textbf{Abernathy-Close L}, \textbf{Perlman NS},
\textbf{Penkevich A}, \textbf{Standke A}, \textbf{Keidan M},
\textbf{Vendrov KC}, \textbf{Bergin IL}, \textbf{Young VB}, \textbf{Rao
K}. 2020. Systemic inflammatory mediators are effective biomarkers for
predicting adverse outcomes in clostridioides difficile infection. mBio
\textbf{11}.
doi:\href{https://doi.org/10.1128/mbio.00180-20}{10.1128/mbio.00180-20}.

\leavevmode\hypertarget{ref-Jenior2018}{}%
30. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2018. \emph{Clostridium difficile} alters the
structure and metabolism of distinct cecal microbiomes during initial
infection to promote sustained colonization. mSphere \textbf{3}.
doi:\href{https://doi.org/10.1128/msphere.00261-18}{10.1128/msphere.00261-18}.

\leavevmode\hypertarget{ref-Vornhagen2020}{}%
31. \textbf{Vornhagen J}, \textbf{Bassis CM}, \textbf{Ramakrishnan S},
\textbf{Hein R}, \textbf{Mason S}, \textbf{Bergman Y}, \textbf{Sunshine
N}, \textbf{Fan Y}, \textbf{Timp W}, \textbf{Schatz MC}, \textbf{Young
VB}, \textbf{Simner PJ}, \textbf{Bachman MA}. 2020. A plasmid locus
associated with klebsiella clinical infections encodes a
microbiome-dependent gut fitness factor.
doi:\href{https://doi.org/10.1101/2020.02.26.963322}{10.1101/2020.02.26.963322}.

\leavevmode\hypertarget{ref-Dieterle2018}{}%
32. \textbf{Dieterle MG}, \textbf{Rao K}, \textbf{Young VB}. 2018. Novel
therapies and preventative strategies for primary and
recurrentClostridium difficileinfections. Annals of the New York Academy
of Sciences \textbf{1435}:110--138.
doi:\href{https://doi.org/10.1111/nyas.13958}{10.1111/nyas.13958}.

\leavevmode\hypertarget{ref-Theriot2011}{}%
33. \textbf{Theriot CM}, \textbf{Koumpouras CC}, \textbf{Carlson PE},
\textbf{Bergin II}, \textbf{Aronoff DM}, \textbf{Young VB}. 2011.
Cefoperazone-treated mice as an experimental platform to assess
differential virulence ofClostridium difficilestrains. Gut Microbes
\textbf{2}:326--334.
doi:\href{https://doi.org/10.4161/gmic.19142}{10.4161/gmic.19142}.

\end{document}