### 2021-04-30 submission/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\vspace{35mm}

\hypertarget{an-osmotic-laxative-renders-mice-susceptible-to-prolonged-clostridioides-difficile-colonization-and-hinders-clearance}{%
\section{\texorpdfstring{An osmotic laxative renders mice susceptible to
prolonged \emph{Clostridioides difficile} colonization and hinders
clearance}{An osmotic laxative renders mice susceptible to prolonged Clostridioides difficile colonization and hinders clearance}}\label{an-osmotic-laxative-renders-mice-susceptible-to-prolonged-clostridioides-difficile-colonization-and-hinders-clearance}}

\vspace{35mm}

Sarah Tomkovich\textsuperscript{1}, Ana Taylor\textsuperscript{1}, Jacob
King\textsuperscript{1}, Joanna Colovas\textsuperscript{1}, Lucas
Bishop\textsuperscript{1}, Kathryn McBride\textsuperscript{1}, Sonya
Royzenblat\textsuperscript{1}, Nicholas A. Lesniak\textsuperscript{1},
Ingrid L. Bergin\textsuperscript{2}, Patrick D.
Schloss\textsuperscript{1\(\dagger\)}

\vspace{40mm}

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:pschloss@umich.edu}{\nolinkurl{pschloss@umich.edu}}

1. Department of Microbiology and Immunology, University of Michigan,
Ann Arbor, MI, USA

2. The Unit for Laboratory Animal Medicine, University of Michigan, Ann
Arbor, MI, USA

\newpage
\linenumbers

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

(Modify depending on target journal, currently abstract submitted to
World Microbe Forum) Antibiotics are a major risk factor for
\emph{Clostridioides difficile} infections (CDIs) because of their
impact on the intestinal microbiome. However, non-antibiotic medications
such as the ubiquitous osmotic laxative polyethylene glycol (PEG) 3350,
also alter the microbiota, but whether PEG impacts CDI susceptibility
and clearance is unclear. To examine how PEG impacts susceptibility, we
treated C57Bl/6 mice with 5-day and 1-day doses of 15\% PEG in the
drinking water and then challenged the mice with \emph{C. difficile} 630
spores. We used clindamycin-treated mice as a control because they
consistently clear \emph{C. difficile} within 10 days post-challenge
(dpc). To examine how PEG treatment impacts clearance, we administered
PEG for 1 day to clindamycin-treated, \emph{C. difficile}-challenged
mice either immediately following challenge or 3 dpc. We collected
longitudinal stool samples to examine \emph{C. difficile} levels in the
stool via anaerobic culture and profiled the microbiota by 16S rRNA
sequencing. PEG treatment alone was sufficient to render mice
susceptible to CDI and 5-day PEG-treated mice remain colonized for up to
30 dpc. Additionally, 5-day PEG treated mice remained susceptible to CDI
10-days post treatment. In contrast, 1-day PEG treated mice were
transiently colonized, clearing \emph{C. difficile} within 7 dpc.
Although 5-day PEG-treated mice exhibited prolonged \emph{C. difficile}
colonization, we saw no difference in histological inflammation between
PEG- and clindamycin-treated mice. Additionally, administering PEG to
mice after \emph{C. difficile} challenge prolonged colonization up to 30
dpc in mice that received PEG immediately after challenge and 15 dpc in
mice that received PEG 3 dpc. When we examined microbiota composition
across our different treatment groups, we found decreased richness in
the PEG-treated mice that exhibited prolonged \emph{C. difficile}
colonization. Importantly, there were increased Bacteroides and
Enterobacteriaceae and decreased Lachnospiraceae and Oscillibacter in
most of the PEG-treated mice with prolonged \emph{C. difficile}
colonization. Our findings suggest the osmotic laxative PEG 3350 alters
the mouse microbiota and disrupts colonization resistance to \emph{C.
difficile}, as well as clearance in mice with a CDI. Considering that
most hospitals recommend not performing \emph{C. difficile} testing on
patients taking laxatives and laxatives are used when administering
fecal microbiota transplants via colonoscopy to patients with recurrent
CDIs, further studies are needed to evaluate if laxatives impact human
microbiota colonization resistance.

\newpage

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Antibiotics are a major risk factor for \emph{Clostridioides difficile}
infections (CDIs) because they disrupt microbiota colonization
resistance (1). However, antibiotics are not the only types of
medications that disrupt the microbiota (2--4). Although, other
medications (proton pump inhibitors, osmotic laxatives, antimotility
agents, and opioids) have been implicated as risk or protective factors
for CDIs through epidemiological studies, whether the association is due
to their impact on the microbiome is still unclear (5--9).

Many of the non-antibiotic medications associated with CDIs are known to
modulate gastrointestinal motility leading to either increased or
decreased colonic transit time, which in turn also strongly impacts
microbiota composition and function (10, 11). Stool consistency often
serves as an approximation of intestinal motility (10). Our group has
shown that when \emph{C. difficile} negative controls are separated into
two groups based on stool consistency, there are shared microbiota
features such as lower alpha diversity in samples from CDI patients and
control patients with diarrhea compared to control samples that were
\emph{C. difficile} negative with non-diarrheal consistency (12). These
results led to a hypothesis that bacterial communities from patients
experiencing diarrhea are susceptible to developing CDIs.

Osmotic laxatives can lead to diarrhea depending on the administered
dose and temporarily disrupt the human intestinal microbiota (13). The
ubiquitous osmotic laxative, polyethylene glycol (PEG) 3350 is found in
Miralax, Nulytely, and Golytely and is also commonly used as bowel
preparation for colonoscopies. Interestingly, previous studies have
shown that treating mice with PEG alone altered microbiota composition,
reduced acetate and butyrate production, altered the mucus barrier, and
rendered the mice susceptible to \emph{C. difficile} infection (14--17).
The mucus barrier is thought to mediate protection from \emph{C.
difficile} infections by protecting intestinal epithelial cells from the
toxins produced by \emph{C. difficile} (18, 19). Whether laxative
administration results in more severe CDIs in mice and how long mice
remain colonized with \emph{C. difficile} after challenge is unclear.

Beyond susceptibility, PEG is also relevant in the context of treating
recurrent CDIs via fecal microbiota transplant (FMT) where a healthy
microbiota is administered to the patient to restore colonization
resistance. For FMTs that are delivered via colonoscopy, patients
typically undergo bowel preparation by taking an osmotic laxative prior
to the procedure. Many of the FMT studies to date rationalize the use of
laxatives (20--22) based on a 1996 case study with 2 pediatric patients
where the authors suggested in the discussion that the laxative may help
flush \emph{C. difficile} spores and toxins from the intestine (23).

Our group has used C57BL/6 mice to characterize how antibiotics disrupt
the microbiota and influence \emph{C. difficile} susceptibility and
clearance (24--26). Although, two groups have now shown PEG treatment
alone renders mice susceptible to \emph{C. difficile} (15, 17), these
studies have raised additional questions regarding the dynamics and
severity of infection as well as the role of laxative treatment in
\emph{C. difficile} clearance. Addressing these questions will better
inform how we think about laxatives in the context of CDIs. Here, we
characterized how long PEG-treated mice remain susceptible, whether PEG
treatment results in more sustained \emph{C. difficile} colonization and
severe CDI and than mice treated with clindamycin, and if PEG treatment
after colonization can promote \emph{C. difficile} clearance.

\hypertarget{results}{%
\subsection{Results}\label{results}}

\textbf{5-day laxative treatment led to prolonged \emph{C. difficile}
colonization in mice.} Building off of previous work that showed
treating mice with the osmotic laxative, PEG 3350, rendered mice
susceptible to \emph{C. difficile} colonization (15, 17), we decided to
test how long \emph{C. difficile} colonization is sustained and how long
PEG-treated mice remain susceptible to \emph{C. difficile}. We compared
three groups of mice treated with PEG 3350 to one group of mice treated
with our standard 10 mg/kg clindamycin treatment, which temporarily
renders the mice susceptible to \emph{C. difficile} colonization, with
mice typically clearing \emph{C. difficile} within 10 days
post-challenge (9, 26). All three groups of PEG-treated mice were
administered a 15\% PEG solution in the drinking water for 5-days, 1.
received no additional treatment, 2. was also treated with clindamycin,
and 3. was allowed to recover for 10 days prior to challenge (Fig. 1A).
PEG treatment resulted in weight loss in all 3 groups of PEG-treated
mice relative to their baseline weights, with the greatest change in
weight observed on the fifth day of PEG treatment. The mice recovered
most of the lost weight by five days after treatment (Fig. 1B). After
either the PEG, clindamycin, or PEG and clindamycin treatment all mice
were challenged with 10\textsuperscript{5} \emph{C. difficile} 630
spores (Fig. 1A). All treatments rendered mice susceptible to \emph{C.
difficile} colonization. However, PEG-treated mice remained colonized
with \emph{C. difficile} at a high level through thirty days
post-challenge (Fig. 1C). In contrast, the clindamycin-treated mice
cleared \emph{C. difficile} within ten days post-challenge (Fig. 1C).
Therefore, PEG treatment led to sustained colonization in contrast to
clindamycin mice that naturally cleared \emph{C. difficile} within ten
days post-challenge.

Notably, we also found PEG-treated mice were still susceptible to
\emph{C. difficile} colonization after a 10-day recovery period,
although \emph{C. difficile} was not detectable in most of the group in
the initial five days post-challenge (Fig. 1C, S1A). One mouse was found
dead on the 6th day post-challenge, presumable due to \emph{C.
difficile} as the bacteria became detecable in that mouse on the 4th day
post-challenge (Fig. S1A, mouse 10). From 8 days post-challenge onward,
the density of \emph{C. difficile} stabilized in the 10-day recovery
group and remained high through 20-30 days post-challenge (Fig. 1C).
Thus, osmotic laxative treatment alone was sufficient to render mice
susceptible to prolonged \emph{C. difficile} colonization and
PEG-treated mice remained susceptible through ten days post-treatment.

\textbf{5-day laxative treatment differentially disrupted the fecal
microbiota compared to clindamycin treatment.} Since osmotic laxatives
and clindamycin have previously been shown to disrupt the murine
microbiota (14--17), we hypothesized the different \emph{C. difficile}
colonization dynamics between mice treated with the osmotic laxative or
clindamycin were due to the two drugs having differential effects on the
microbiota. We profiled the stool microbiota over time by sequencing the
V4 region of the 16S rRNA gene to compare changes across treatment
groups. We found time and treatment group explained half of the observed
variation between fecal communities with most of the remaining variation
explained by interactions between treatment group and other experimental
variables including time, cage effects, and sequencing preparation plate
(PERMANOVA combined R\textsuperscript{2} = 0.95, \emph{P} \textless{}
0.001, Fig. 2A, Data Set S1, Sheet X). Cage effects refer to the
well-documented phenomenon that mice housed in the same cages have
similar microbial communities due to coprophagy (27). We tried to
minimize the impact of cage effects on our experiment by breaking up
cagemates when assigning mice to treatment groups. Importantly, although
we conducted a total of 5 separate experiments, the experiment number
and its interaction with treatment group was not one of the variables
that significantly explained the observed variation in fecal communities
(Data Set S1, Sheet X). None of the treatment groups recovered to their
baseline community structure either 10 or 30 days post-challenge
suggesting other community features besides recovery to baseline were
responsible for the prolonged \emph{C. difficile} colonization in
PEG-treated mice (Fig. 2B).

Since none of the communities completely recovered in the follow-up
period after treatments, we next profiled community diversity and
composition. We examined the alpha diversity dynamics by calculating the
communities' Shannon diversity. Although both clindamycin and PEG
treatments decreased diversity, Shannon diversity was lower in the
groups of mice that received PEG treatment compared to those that
received clindamycin through thirty days post-challenge (Fig. 2C). We
next examined the bacterial genera that shifted after PEG treatment by
comparing the baseline samples of mice treated with only PEG to samples
from the same mice one day post-PEG-treatment. We found 18 genera that
were altered by PEG treatment (Data Set S1, Sheet X). The majority of
the bacterial relative abundances decreased after the PEG treatment, but
the relative abundance among members of the \emph{Enterobacteriaceae}
and \emph{Bacteroides} increased. The increase in \emph{Bacteroides}
relative abundance was unique to PEG treated mice, as the
\emph{Bacteroides} relative abundance actually decreased in clindamycin
treated mice (Fig. 2D). Finally, we identified the genera whose relative
abundance differed across treatment groups over multiple time points. Of
the 33 genera that were different between treatment groups, 24 genera
were different over multiple time points (Fig. 2E, Data Set S1, Sheet
X). Thus, PEG had a significant impact on the fecal microbiota that was
maintained over time and was distinct from clindamycin treatment.

Interestingly, \emph{C. difficile} was not immediately detectable in the
stools of the PEG-treated mice that were allowed to recover for 10 days
prior to challenge. We decided to examine if there were genera that
changed during the post-challenge period when the group median \emph{C.
difficile} shifted from undetectable at 1 day post-challenge to
detectable in the stool samples with the density stabilizing around 8
days post-challenge (Fig. S1A). We found no bacteria were significantly
different over the two time periods after multiple hypothesis correction
(Data Set S1, Sheet X). However, there was also wide variation between
individual mice regarding when \emph{C. difficile} became detectable
(Fig. S1A) as well as the relative abundances of bacterial genera in the
communities (Fig. S1B). For example, two mice had a high relative
abundance of \emph{Enterobacteriaceae} throughout the post-challenge
period and this corresponded to mouse 10, which died on the sixth day
post-challenge and mouse 11, where \emph{C. difficile} was present at a
high density from the 4th day post-challenge onward (Fig. S1B). While we
did not identify a clear signal to explain the delayed appearance of
\emph{C. difficile} in the 5-day PEG mice that were allowed to recover
for 10 days prior to challenge, the delay is striking and could reflect
changes in microbial activity or metabolites that were not examined in
this study.

\textbf{5-day laxative treatment did not promote more severe CDIs
despite altering the mucosal microbiota.} Given the findings from a
previous study that demonstrated PEG treatment disrupts the mucus layer
and alters the immune response in mice (16), we decided to examine the
impact of PEG treatment on the mucosal microbiota and CDI severity. To
evaluate the mucosal microbiota, we sequenced snips of tissue collected
from the cecum, proximal colon, and distal colon. Similar to what was
observed with the stool samples, alpha diversity was lower in the
PEG-treated mice compared to clindamycin treated mice (Fig. 3A). Alpha
diversity continued to increase over time with the PEG-treated mice
collected at 20 and 30 days post-challenge (Fig. 3A, Data Set S1, Sheet
X). Group, time point, and their interactions with other variables
(cage, experiment number, and sample type) explained the majority of the
variation observed in mucosal communities (PERMANOVA combined
R\textsuperscript{2} = 0.83, \emph{P} \textless{} 0.05, Fig. 3B, Data
Set S1, Sheet X). We saw the greatest difference in the relative
abundance of the mucosal microbiota between treatment groups
(clindamycin, 5-day PEG, and 5-day PEG plus clindamycin) at 6 days
post-challenge with 10 genera that were significantly different
(\emph{P} `r format.pval(tissue\_genera\_max\_p, eps = 0.05)) in all
three of the tissue types we collected (cecum, proximal colon, and
distal colon; Fig. S2A, Data Set S1, Sheet X). Interestingly,
\emph{Peptostreptococcaceae} (the family with a sequence that matches
\emph{C. difficile}) was one of the genera that had a significant
difference in relative abundance between treatment groups at 6 days
post-challenge. This population was primarily only present in the 5-day
PEG treatment group of mice and decreased in the proximal and distal
colon tissues over time (Fig. S2B). By 30 days post-challenge, only the
relative abundances of \emph{Bacteroides}, \emph{Clostridiales},
\emph{Firmicutes}, and \emph{Ruminococcaceae} were different between
treatment groups and only in the cecum tissues (Fig. 3C, Fig. 2E, Data
Set S1, Sheet X). Thus, PEG treatment had a significant impact on the
mucosal microbiota and we detected \emph{C. difficile} sequences in the
cecum, proximal colon, and distal colon tissue communities.

Next, we examined the severity of \emph{C. difficile} challenge by
evaluating cecum and colon H\&E stained histopathology (28) and found
there was no difference in cecum and colon scores between clindamycin
and PEG-treated mice that were challenged with \emph{C. difficile} at 4
days post-challenge (Fig. 3D), the time point typically examined in
\emph{C. difficile} 630 challenged mice (29). We also looked at 6 days
post-challenge because that was when there was a large difference in
\emph{C. difficile} density between PEG- and clindamycin-treated mice
(Fig. 1C). Although, there was a slight difference in the colon between
PEG and clindamycin-treated mice, there was not a signifant difference
in the cecum and the overall score was relatively low (1.5-2.5 out of
12, Fig. 3E). Therefore, although PEG treatment had a disruptive effect
on the mucosal microbiota, the impact of \emph{C. difficile} 630
challenge on the cecum and colon was similar between PEG and clindamycin
treated mice.

\textbf{\emph{C. difficile} challenge did not have a synergistic
disruptive effect on the microbiota of PEG-treated mice.} Because
\emph{C. difficile} itself can have an impact on the microbiota (30), we
also sequenced the tissue and stools of mock-challenged clindamycin and
5-day PEG treated mice. Examining the stools of the mock-challenged mice
revealed similar bacterial disruptions as the \emph{C. difficile}
challenged mice (Fig. S3A-C). Similarly, there was no difference between
the tissues of mock and \emph{C. difficile} challenged mice (Fig.
S3D-F). Thus, most of the microbiota alterations we observed in the
PEG-treated mice were a result of the laxative and not an interaction
between the laxative and \emph{C. difficile}.

\textbf{1-day laxative treatment resulted in transient \emph{C.
difficile} colonization and minor microbiota disruption.} Next, we
examined how a shorter osmotic laxative perturbation would impact the
microbiome and susceptibility to \emph{C. difficile}. We administered
either a 1-day PEG treatment, a 1-day PEG treatment with a 1-day
recovery period, or clindamycin to mice before challenging them with
\emph{C. difficile} (Fig. 3A). In contrast to the 5-day PEG treated
mice, the 1-day PEG treated mice were only transiently colonized and
cleared \emph{C. difficile} by 7 days post-challenge (Fig. 3B). The
stool communities of PEG-treated mice were also only transiently
disrupted, with Shannon diversity recovering by 7 days post-challenge
(Fig. 3C-D). We found the relative abundances of 14 genera were impacted
by treatment, but recovered close to baseline levels by 7 days
post-challenge including \emph{Enterobacteriaceae},
\emph{Clostridiales}, \emph{Porpyromonadaceae}, and
\emph{Ruminococcaceae} (Fig. 3E, Data Set S1, Sheet X). These findings
suggest the duration of the PEG treatment was relevant, with shorter
treatments resulting in a transient loss of \emph{C. difficile}
colonization resistance.

\textbf{Post-CDI laxative treatment disrupted clearance in
clindamycin-treated mice regardless of whether an FMT was also
administered.} Since a 1-day PEG treatment resulted in a more mild
microbiota perturbation, we decided to use the 1-day treatment to
examine the hypothesis that PEG helps to flush \emph{C. difficile}
spores from the intestine. To examine the impact of PEG treatment on
\emph{C. difficile} clearance, we treated 4 groups of mice with
clindamycin and then challenged all mice with \emph{C. difficile} before
administering the following treatments: no additional treatment, 1-day
PEG immediately after challenge, and 1-day PEG treatment 3 days after
challenge followed by either administration of an FMT or PBS solution by
oral gavage (Fig. 5A). Contrary to our hypothesis, all groups of mice
that received PEG exhibited prolonged \emph{C. difficile} colonization
(Fig. 5B). Next we examined how post-CDI PEG treatment impacted
microbiota diversity. Alpha diversity was lower in the PEG-treated mice
with the exception of the PEG-treated mice that were administered an FMT
(Fig 5C-D).

We were also interested in exploring whether PEG might help with
engraftment in the context of FMTs. The FMT appeared to partially
restore Shannon diversity but not richness (Fig. 5C-D). Similarly, we
saw some overlap between the communities of mice that received FMT and
the mice treated with only clindamycin after 5 days post-challenge (Fig.
6A). The increase in Shannon diversity suggests that the FMT did have an
impact on the microbiota, despite seeing prolonged \emph{C. difficile}
colonization in the FMT treated mice. However, only the relative
abundances of \emph{Bacteroidales} and \emph{Porphyromonadaceae}
consistently differed between the mice received either an FMT or PBS
gavage (Fig. 6B), suggesting the FMT only restored a couple of genera.
Overall, we found the relative abundances of 24 genera were different
between treatment groups over multiple timepoints. For example, the
relative abundance of \emph{Akkermansia} was increased and the relative
abundances of \emph{Ruminococcaceae}, \emph{Clostridiales},
\emph{Lachnospiraceae}, and \emph{Oscillibacter} were decreased in mice
that received PEG after \emph{C. difficile} challenge relative to
clindamycin treated mice (Fig. 6C). In sum, administering PEG actually
prolonged \emph{C. difficile} colonization, including in mice that
received an FMT, which only restored 2 bacterial genera.

\textbf{Five-day post-challenge community data can predict which mice
that will have prolonged \emph{C. difficile} colonization.} After
identifying bacteria associated with the 5-day, 1-day and post-CDI 1-day
PEG treatments, we decided to examine the bacteria that influenced
prolonged \emph{C. difficile} colonization. We trained 3 types of
machine learning models (random forest, logistic regression, and support
vector machine) with bacterial community data from 5 days post-challenge
to predict whether the mice were still colonized with \emph{C.
difficile} 10 days post-challenge. We chose 5 days post-challenge
because that was the earliest time point where we would see a treatment
effect in the mice that were given 1-day PEG treatment three days after
challenge. The random forest model had the highest performance (AUROC =
0.90, Data Set S1, Sheet X), so we next performed permutation importance
to examine the bacteria that were the top contributors to the random
forest model predicting prolonged \emph{C. difficile }colonization. We
selected the top 10 bacteria contributing to our models performance
(Fig. 7A) and examined their relative abundance at 5 days
post-challenge, the time point used to predict \emph{C. difficile}
colonization status on day 10 (Fig. 7B). Next, we focused on the 5
genera that had a greater than 1 \% relative abundance in either the
cleared or colonized mice and examined how the bacteria changed over
time. We found \emph{Enterobacteriaceae} and \emph{Bacteroides} tended
to have a higher relative abundance, the relative abundance of
\emph{Akkermansia} was initially decreased and then increased, and
\emph{Porphyromonadaceae} and \emph{Lachnospiraceae} had a lower
relative abundance in the mice with prolonged colonization compared to
the mice that cleared \emph{C. difficile} (Fig. 7C). Together these
results suggest a combination of low and high abundance bacterial genera
influence the prolonged colonization observed in 5-day PEG and post-CDI
PEG treated mice.

Previous work examining the impact of PEG on the murine microbiota found
that PEG treatment resulted in the permanent loss of \emph{Muribaculum
intestinale} (16)., which is classified as \emph{Porphyromonadaceae} by
the Ribosomal Database Project (RDP) database (31).
\emph{Porphyromonadaceae} was a top feature in the random forest model
predicting prolonged \emph{C. difficile} colonization. We identified 4
OTUs that had 92-96\% identity to \emph{Muribaculum intestinale} and
examined their abundance in mice that either cleared or were still
colonized with \emph{C. difficile} at 10 days post-challenge. While all
of the OTUs, were decreased by PEG and clindamycin treatment, there was
some recovery in the mice that cleared (Fig. S4A). We also examined
other \emph{Porphyromonadaceae} and \emph{Lachnospiraceae} OTUs since
these were the 2 genera that were important to our classification model
and contained multiple OTUs that were different at 5 days post-challenge
between mice that either cleared or remained colonized with \emph{C.
difficile} by 10 days post-challenge (Data Set S1, sheet X). While
individual \emph{Porphyromonadaceae} and \emph{Lachnospiraceae} OTUs
tended to be more abundant in the mice that clear \emph{C. difficile}
relative to the mice that exhibit prolonged colonization (Fig. S4B-C),
there is no single OTU that fits the pattern we observed at the genus
level (Fig. 7C), suggesting multiple \emph{Porphyromonadaceae} and
\emph{Lachnospiraceae} OTUs influenced \emph{C. difficile} clearance.
Overall, our results suggest that specific bacterial community
differences explain the prolonged \emph{C. difficile} colonization we
observed in 5-day PEG and post-CDI 1-day PEG treated mice.

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

\begin{itemize}
\item
  Summary of major findings (Fig. 8A)
\item
  Discussion of prolonged persistence. \emph{C. difficile} sequences
  detected in tissue samples. Association with mucin-degrading bacteria
  suggested by recent papers.
\item
  Discuss why we might not have observed more severe histology in PEG
  mice relative to clindamycin-treated mice

  \begin{itemize}
  \tightlist
  \item
    Antibiotics may also impact mucus layer
  \item
    Strain of bacteria used
  \end{itemize}
\item
  Protective bacteria missing in PEG-treated mice
\item
  Discuss what these findings might mean for human patients (Fig. 8B)

  \begin{itemize}
  \tightlist
  \item
    What's known regarding laxatives and susceptibility to CDIs
  \item
    Clinical trial of PEG, results never posted (32)
  \item
    Relevance to human FMTs? Unclear what the best administration route
    is because there have been no studies designed to evaluate the best
    administration route for FMTs.
  \end{itemize}
\end{itemize}

\hypertarget{conclusions}{%
\subsection{Conclusions}\label{conclusions}}

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

We thank members of the Schloss lab for feedback on planning the
experiments and data presentation. We thank Andrew Henry for help with
media preparation and bacterial culture and the Microbiology and
Immunology department's postdoc association writing group members for
their feedback on manuscript drafts. We also thank the Unit for
Laboratory Animal Medicine at the University of Michigan for maintaining
our mouse colony and providing the institutional support for our mouse
experiments. Finally, we thank Kwi Kim, Austin Campbell, and Kimberly
Vendrov for their help in maintaining the Schloss lab's anaerobic
chamber. This work was supported by the National Institutes of Health
(U01AI124255). ST was supported by the Michigan Institute for Clincial
and Health Research Postdoctoral Translation Scholars Program
(UL1TR002240 from the National Center for Advancing Translational
Sciences).

\hypertarget{materials-and-methods}{%
\subsection{Materials and Methods}\label{materials-and-methods}}

\begin{itemize}
\tightlist
\item
  Histopathology (33)
\end{itemize}

\newpage

\includegraphics{figure_1.pdf}

\textbf{Figure 1. 5-day PEG treatment prolongs susceptibility and mice
become persistently colonized with \emph{C. difficile}.} A. Setup of the
experimental time line for experiments with 5-day PEG treated mice
consisting of 4 treatment groups. 1. Clindamycin was administered at 10
mg/kg by intraperitoneal injection. 2. 15\% PEG 3350 was administered in
the drinking water for five days. 3. 5-day PEG plus clindamycin
treatment. 4. 5-day PEG plus 10-day recovery treatment. All treatment
groups were then challenged with 10\textsuperscript{5} \emph{C.
difficile} 630 spores. A subset of mice were euthanized on either 4 or 6
days post-challenge and tissues were collected for histopathology
analysis, the remaining mice were followed through 20 or 30 days
post-challenge. B. Weight change from baseline weight in groups after
treatment with PEG and/or clindamycin, followed by \emph{C. difficile}
challenge. C. \emph{C. difficile} CFU/gram stool measured over time (N =
10-59 mice per time point) via serial dilutions. The black line
represents the limit of detection for the first serial dilution. CFU
quantification data was not available for each mouse due to stool
sampling difficulties (particularly the day the mice came off of the PEG
treatment) or early deaths. For B-C, lines represent the median for each
treatment group and circles represent samples from individual mice.
Asterisks indicate time points where the weight change or CFU/g was
significantly different between groups by the Kruskal-Wallis test with
Benjamini-Hochberg correction for testing multiple time points. The data
presented are from a total of 5 separate experiments. \newpage

\includegraphics{figure_2.pdf} \textbf{Figure 2. 5-day PEG treatment
disrupts the stool microbiota for a longer amount of time compared to
clindamycin-treated mice.} A. Principal Coordinate analysis (PCoA) of
Bray-Curtis distances from stool samples collected throughout the
experiment. For A and C, each circle represents a sample from an
individual mouse and the transparency of the symbol corresponds to the
day post-challenge. B. Bray-Curtis distances of stool samples collected
on either day 10 or 30 post-challenge relative to the baseline sample
collected for each mouse (before any drug treatments were administered).
C. Shannon diversity in stool communities over time. The line indicates
the median value for each treatment group. The colors of the symbols and
lines represent the four treatment groups. D. 14 of the 33 genera
affected by PEG treatment (Data Set S1, sheet X). The symbols represent
the median relative abundance for a treatment group at either baseline
(open circle) or 1-day post treatment (closed circle). Data from the
5-day PEG and 5-day PEG plus 10-day recovery groups were analyzed by
paired Wilcoxan signed-rank test with Benjamini-Hochberg correction for
testing all identified genera. The clindamycin and 5-day PEG plus
clindamycin treatment groups are shown for comparison. E. 6 of the 24
genera that were significantly different between the four treatment
groups over multiple time points. Differences between treatment groups
were identified by Kruskal-Wallis test with Benjamini-Hochberg
correction for testing all identified genera. The gray vertical line (D)
and horizontal vertical lines (E) indicate the limit of detection.
\newpage

\includegraphics{figure_3.pdf} \textbf{Figure 3. 5-day PEG treatment
does not result in more severe CDIs, although mucosal microbiota is
altered.} A. Shannon diversity in cecum communities over time. The line
indicates the median value for each treatment group. The colors of the
symbols and lines represent the four treatment groups. A similar pattern
was observed with the proximal and distal colon communities (Data Set
S1, sheet X-X). B. PCoA of Bray-Curtis distances from mucosal samples
collected throughout the experiment. Circles, triangles, and squares
indicate cecum, proximal colon, and distal colon communities,
respectively. For A-B, transparency of the symbol corresponds to the day
post-challenge that the sample was collected. C. The median relative
abundance of the 4 genera that were significantly different between the
cecum communities of different treatment groups on day 6 and day 30
(Data Set S1, sheet X). The gray vertical line indicate the limit of
detection. D-E. The histopathology summary scores from cecum and colon
H\&E stained slides. The summary score is the total score based on
evaluation of edema, cellular infiltration, and inflammation. Each
category is given a score ranging from 0-4, thus the maximum possible
summary score is 12. The tissue for histology was collected at either 4
(D) or 6 (E) days post-challenge with the exception that one set of
5-day PEG treated mock-challenged mice were collected on day 0
post-challenge (first set of open circles in D). Histology data were
analyzed with the Kruskal-Wallis test followed by pairwise Wilcoxon
comparisons with Benjamini-Hochberg correction. \newpage

\includegraphics{figure_4.pdf} \textbf{Figure 4. 1-day PEG treatment
renders mice susceptible to transient \emph{C. difficile} colonization.}
A. Setup of the experimental time line for the 1-day PEG treated mice
consisting of 3 treatment groups. 1. Clindamycin was administered at 10
mg/kg by intraperitoneal injection. 2. 15\% PEG 3350 was administered in
the drinking water for 1 day. 3. 1-day PEG plus 1-day recovery. The
three treatment groups were then challenged with 10\textsuperscript{5}
\emph{C. difficile} 630 spores. B. \emph{C. difficile} CFU/gram stool
measured over time (N = 12-18 mice per time point) via several
dilutions. The black dotted line represents the limit of detection for
the first serial dilution. Asterisks indicate time points where the
CFU/gram was significantly different between treatment groups by
Kruskall-Wallis test with Benjamini-Hochberg correction for testing
multiple time points. C. PCoA of Bray-Curtis distances from stool
communities collected from the three treatment groups over time (day:
R\textsuperscript{2} = 0.43; group: R\textsuperscript{2} = 0.19). D.
Shannon diversity in stool communities over time with colored lines
representing the median value for each treatment group. For B-D, each
symbol represents a sample from an individual mouse and symbol
transparency corresponds to the day post-challenge that the sample was
collected. E. Median relative abundances per treatment group for 6 out
of the 14 genera that were affected by treatment, but recovered close to
baseline levels by 7 days post-challenge (Fig. 3E, Data Set S1, Sheet
X). Stool samples from either baseline and day 1 or baseline and day 7
were analyzed by paired Wilcoxan signed-rank test with
Benjamini-Hochberg correction for testing all identified genera. The
gray horizontal line represents the limit of detection.

\includegraphics{figure_5.pdf} \textbf{Figure 5. 1-day PEG treatment
post \emph{C. difficile} challenge prolongs colonization regardless of
whether an FMT is also administered.} A. Setup off the experimental time
line for experiments with post-CDI PEG treated mice. There were a total
of 4 different treatment groups. All mice were administered 10 mg/kg
clindamycin intraperitoneally (IP) 1 day before challenge with
10\textsuperscript{3-5} \emph{C. difficile} 630 spores. 1. Did not
receive any additional treatment. 2. Immediately after \emph{C.
difficile} challenge, mice received 15\% PEG 3350 in the drinking water
for 1 day. 3-4. 3-days after challenge, mice received 1-day PEG
treatment and then received either a fecal microbiota transplant (3) or
PBS (4) solution by oral gavage. Mice were followed through 15-30 days
post-challenge (only the post-CDI 1-day PEG group was followed through
30 days post-challenge). B. CFU/g of \emph{C. difficile} stool measured
over time via serial dilutions. The black line represents the limit of
detection for the first serial dilution. C-D. Shannon diversity (C) and
richness (D) in stool communities over time. B-D. Each symbol represents
a stool sample from an individual mouse with the lines representing the
median value for each treatment group. The transparancy of the symbol
corresponds to the day post-challenge. Asterisks indicate time points
with significant differences between groups by a Kruskall-Wallis test
with a Benjamini-Hochberg correction for testing multiple times points.
Colored rectangles indicates the 1-day PEG treatment period for
applicable groups. The data presented are from a total of 3 separate
experiments. \newpage

\includegraphics{figure_6.pdf} \textbf{Figure 6. For 1-day PEG treatment
post \emph{C. difficile} challenge mice that also receive an FMT only
some bacterial genera were restored.} A. PCoA of Bray-Curtis distances
from stool samples collected over time as well as the FMT solution that
was administered to one treatment group. Each circle represents an
individual sample, the transparency of the circle corresponds to day
post-challenge as shown in Fig. 6C-D. B. Median relative abundances of 2
genera that were significantly different over multiple time points in
mice that were administered either FMT or PBS solution via gavage C.
Median relative abundances of the top 6 out of 24 genera that were
significant over multiple timepoints, plotted over time (Data Set S1,
Sheet X). For B-C, colored rectangles indicates the 1-day PEG treatment
period for applicable groups. Gray horizontal lines represent the limit
of detection.

\newpage

\includegraphics{figure_7.pdf} \textbf{Figure 7. Specific microbiota
features associated with prolonged \emph{C. difficile} colonization in
PEG treated mice.} A. Top ten bacteria that contributed to the random
forest model trained on five day post-challenge community relative
abundance data to predict wehther mice would still be colonized with
\emph{C. difficile} 10 days post-challenge. The median (point) and
interquantile range (lines) change in AUROC when the bacteria is left
out of the model is shown. B. The median relative abundances at 5 days
post-challenge of the top ten bacteria that contributed to the random
forest classification model. Color indicates whether the mice were still
colonized with \emph{C. difficile} 10 days post-challenge and the black
horizontal line represents the median relative abundance. Each symbol
represents a stool sample from an individual mouse and the shape of the
symbol indicates whether the PEG-treated mice recieved a 5-day (Fig.
1-3), 1-day (Fig. 4) or post-CDI PEG (Fig. 5-6) treatment. C. The median
relative abundances of the 5 genera with greater than 1\% median
relative abundance in the stool community over time. For B-C, the gray
horizontal line represents the limit of detection. \newpage

\includegraphics{figure_8.pdf} \textbf{Figure 8. Schematic summarizing
findings.} A. The gut microbiota of our C57Bl/6 mice is restant to
\emph{C. difficle} but treatment with either the antibiotic,
clindamycin, or the osmotic laxative, PEG 3350 renders the mice
susceptible to \emph{C. difficile} colonization. Recovery of
colonization resistance in clindamycin-treated mice is relatively
straightforward and the mice clear \emph{C.difficile} within 10 days
post-challenge. However, for mice that recieved a 5-day PEG treatment or
a 1-day PEG treatment after \emph{C. difficile} challenge recovery of
colonization resistance is more uncertain because mice were still
colonized with \emph{C. difficile} 30 days post-challenge in the case of
several PEG treatments. We found increased \emph{Porphyromonadaceae} and
\emph{Lachnospiraceae} were associated with recovery of colonization
resistance, while increased \emph{Enterobacteriaceae} and
\emph{Bacteroides} were associated with prolonged \emph{C. difficle}
colonization. B.Considering that most hospitals recommend not performing
\emph{C. difficile} testing on patients taking laxatives and laxatives
are used when administering fecal microbiota transplants via colonoscopy
to patients with recurrent CDIs. further studies are needed to evaluate
if laxatives impact human microbiota colonization resistance. Further
studies are needed to understand the impact of osmotic laxatives on
\emph{C. difficile} colonization resistance and clearance in human
patients. \newpage

\includegraphics{figure_S1.pdf} \textbf{Figure S1. Microbiota dynamics
post-challenge in the 5-day PEG treatment plus 10-day recovery mice.} A.
\emph{C. difficile} CFU/g over time in the stool samples collected from
5-day PEG treatment plus 10-day recovery mice. Same data presented in
Fig. 1C, but the data for the other 3 treatment groups have been
removed. B. Median relative abundances of 8 bacterial genera from day 0
post-challenge onward from the 10-day recovery mice. We analyzed samples
from day 0 and day 8 post-challenge, which represented the the time
points where mice were challenged with \emph{C. difficile} and when the
median relative \emph{C. difficile} CFU stabilized for the group using
the paired Wilcoxan signed-rank test, but no genera were signifcant
after Benjamini-Hochberg correction. \newpage

\includegraphics{figure_S2.pdf} \textbf{Figure S2. PEG treatment still
has a large impact on the mucosal microbiota 6 days post-challenge} A.
The\\
\newpage

\includegraphics{figure_S3.pdf} \textbf{Figure S3. \emph{C. difficile}
challenge does not enhance the disruptive effect of PEG on the
microbiota.} A. \newpage

\includegraphics{figure_S4.pdf} \textbf{Figure S4. Specific OTUs
associated with clearance by 10 days post-challenge that are mostly
absent in mice with prolonged \emph{C. difficile} colonization. Ex.
\emph{Muribaculum intestinale}.} A. \newpage

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Britton2014}{}%
1. \textbf{Britton RA}, \textbf{Young VB}. 2014. Role of the intestinal
microbiota in resistance to colonization by clostridium difficile.
Gastroenterology \textbf{146}:1547--1553.
doi:\href{https://doi.org/10.1053/j.gastro.2014.01.059}{10.1053/j.gastro.2014.01.059}.

\leavevmode\hypertarget{ref-Maier2018}{}%
2. \textbf{Maier L}, \textbf{Pruteanu M}, \textbf{Kuhn M},
\textbf{Zeller G}, \textbf{Telzerow A}, \textbf{Anderson EE},
\textbf{Brochado AR}, \textbf{Fernandez KC}, \textbf{Dose H},
\textbf{Mori H}, \textbf{Patil KR}, \textbf{Bork P}, \textbf{Typas A}.
2018. Extensive impact of non-antibiotic drugs on human gut bacteria.
Nature \textbf{555}:623--628.
doi:\href{https://doi.org/10.1038/nature25979}{10.1038/nature25979}.

\leavevmode\hypertarget{ref-LeBastard2017}{}%
3. \textbf{Bastard QL}, \textbf{Al-Ghalith GA}, \textbf{Grégoire M},
\textbf{Chapelet G}, \textbf{Javaudin F}, \textbf{Dailly E},
\textbf{Batard E}, \textbf{Knights D}, \textbf{Montassier E}. 2017.
Systematic review: Human gut dysbiosis induced by non-antibiotic
prescription medications. Alimentary Pharmacology \& Therapeutics
\textbf{47}:332--345.
doi:\href{https://doi.org/10.1111/apt.14451}{10.1111/apt.14451}.

\leavevmode\hypertarget{ref-VichVila2020}{}%
4. \textbf{Vila AV}, \textbf{Collij V}, \textbf{Sanna S}, \textbf{Sinha
T}, \textbf{Imhann F}, \textbf{Bourgonje AR}, \textbf{Mujagic Z},
\textbf{Jonkers DMAE}, \textbf{Masclee AAM}, \textbf{Fu J},
\textbf{Kurilshikov A}, \textbf{Wijmenga C}, \textbf{Zhernakova A},
\textbf{Weersma RK}. 2020. Impact of commonly used drugs on the
composition and metabolic function of the gut microbiota. Nature
Communications \textbf{11}.
doi:\href{https://doi.org/10.1038/s41467-019-14177-z}{10.1038/s41467-019-14177-z}.

\leavevmode\hypertarget{ref-Oh2018}{}%
5. \textbf{Oh J}, \textbf{Makar M}, \textbf{Fusco C}, \textbf{McCaffrey
R}, \textbf{Rao K}, \textbf{Ryan EE}, \textbf{Washer L}, \textbf{West
LR}, \textbf{Young VB}, \textbf{Guttag J}, \textbf{Hooper DC},
\textbf{Shenoy ES}, \textbf{Wiens J}. 2018. A generalizable, data-driven
approach to predict daily risk ofClostridium difficileInfection at two
large academic health centers. Infection Control \& Hospital
Epidemiology \textbf{39}:425--433.
doi:\href{https://doi.org/10.1017/ice.2018.16}{10.1017/ice.2018.16}.

\leavevmode\hypertarget{ref-Mora2012}{}%
6. \textbf{Mora AL}, \textbf{Salazar M}, \textbf{Pablo-Caeiro J},
\textbf{Frost CP}, \textbf{Yadav Y}, \textbf{DuPont HL}, \textbf{Garey
KW}. 2012. Moderate to high use of opioid analgesics are associated with
an increased risk of clostridium difficile infection. The American
Journal of the Medical Sciences \textbf{343}:277--280.
doi:\href{https://doi.org/10.1097/maj.0b013e31822f42eb}{10.1097/maj.0b013e31822f42eb}.

\leavevmode\hypertarget{ref-Nehra2018}{}%
7. \textbf{Nehra AK}, \textbf{Alexander JA}, \textbf{Loftus CG},
\textbf{Nehra V}. 2018. Proton pump inhibitors: Review of emerging
concerns. Mayo Clinic Proceedings \textbf{93}:240--246.
doi:\href{https://doi.org/10.1016/j.mayocp.2017.10.022}{10.1016/j.mayocp.2017.10.022}.

\leavevmode\hypertarget{ref-Krishna2013}{}%
8. \textbf{Krishna SG}, \textbf{Zhao W}, \textbf{Apewokin SK},
\textbf{Krishna K}, \textbf{Chepyala P}, \textbf{Anaissie EJ}. 2013.
Risk factors, preemptive therapy, and antiperistaltic agents
forClostridium difficileinfection in cancer patients. Transplant
Infectious Disease n/a--n/a.
doi:\href{https://doi.org/10.1111/tid.12112}{10.1111/tid.12112}.

\leavevmode\hypertarget{ref-Tomkovich2019}{}%
9. \textbf{Tomkovich S}, \textbf{Lesniak NA}, \textbf{Li Y},
\textbf{Bishop L}, \textbf{Fitzgerald MJ}, \textbf{Schloss PD}. 2019.
The proton pump inhibitor omeprazole does not promote
\emph{Clostridioides difficile} colonization in a murine model. mSphere
\textbf{4}.
doi:\href{https://doi.org/10.1128/msphere.00693-19}{10.1128/msphere.00693-19}.

\leavevmode\hypertarget{ref-Vandeputte2015}{}%
10. \textbf{Vandeputte D}, \textbf{Falony G}, \textbf{Vieira-Silva S},
\textbf{Tito RY}, \textbf{Joossens M}, \textbf{Raes J}. 2015. Stool
consistency is strongly associated with gut microbiota richness and
composition, enterotypes and bacterial growth rates. Gut
\textbf{65}:57--62.
doi:\href{https://doi.org/10.1136/gutjnl-2015-309618}{10.1136/gutjnl-2015-309618}.

\leavevmode\hypertarget{ref-VujkovicCvijin2020}{}%
11. \textbf{Vujkovic-Cvijin I}, \textbf{Sklar J}, \textbf{Jiang L},
\textbf{Natarajan L}, \textbf{Knight R}, \textbf{Belkaid Y}. 2020. Host
variables confound gut microbiota studies of human disease. Nature
\textbf{587}:448--454.
doi:\href{https://doi.org/10.1038/s41586-020-2881-9}{10.1038/s41586-020-2881-9}.

\leavevmode\hypertarget{ref-Schubert2014}{}%
12. \textbf{Schubert AM}, \textbf{Rogers MAM}, \textbf{Ring C},
\textbf{Mogle J}, \textbf{Petrosino JP}, \textbf{Young VB},
\textbf{Aronoff DM}, \textbf{Schloss PD}. 2014. Microbiome data
distinguish patients with clostridium difficile infection and non-c.
Difficile-associated diarrhea from healthy controls. mBio \textbf{5}.
doi:\href{https://doi.org/10.1128/mbio.01021-14}{10.1128/mbio.01021-14}.

\leavevmode\hypertarget{ref-Nagata2019}{}%
13. \textbf{Nagata N}, \textbf{Tohya M}, \textbf{Fukuda S}, \textbf{Suda
W}, \textbf{Nishijima S}, \textbf{Takeuchi F}, \textbf{Ohsugi M},
\textbf{Tsujimoto T}, \textbf{Nakamura T}, \textbf{Shimomura A},
\textbf{Yanagisawa N}, \textbf{Hisada Y}, \textbf{Watanabe K},
\textbf{Imbe K}, \textbf{Akiyama J}, \textbf{Mizokami M},
\textbf{Miyoshi-Akiyama T}, \textbf{Uemura N}, \textbf{Hattori M}. 2019.
Effects of bowel preparation on the human gut microbiome and metabolome.
Scientific Reports \textbf{9}.
doi:\href{https://doi.org/10.1038/s41598-019-40182-9}{10.1038/s41598-019-40182-9}.

\leavevmode\hypertarget{ref-Kashyap2013}{}%
14. \textbf{Kashyap PC}, \textbf{Marcobal A}, \textbf{Ursell LK},
\textbf{Larauche M}, \textbf{Duboc H}, \textbf{Earle KA},
\textbf{Sonnenburg ED}, \textbf{Ferreyra JA}, \textbf{Higginbottom SK},
\textbf{Million M}, \textbf{Tache Y}, \textbf{Pasricha PJ},
\textbf{Knight R}, \textbf{Farrugia G}, \textbf{Sonnenburg JL}. 2013.
Complex interactions among diet, gastrointestinal transit, and gut
microbiota in humanized mice. Gastroenterology \textbf{144}:967--977.
doi:\href{https://doi.org/10.1053/j.gastro.2013.01.047}{10.1053/j.gastro.2013.01.047}.

\leavevmode\hypertarget{ref-Ferreyra2014}{}%
15. \textbf{Ferreyra JA}, \textbf{Wu KJ}, \textbf{Hryckowian AJ},
\textbf{Bouley DM}, \textbf{Weimer BC}, \textbf{Sonnenburg JL}. 2014.
Gut microbiota-produced succinate promotes c.~difficile infection after
antibiotic treatment or motility disturbance. Cell Host \& Microbe
\textbf{16}:770--777.
doi:\href{https://doi.org/10.1016/j.chom.2014.11.003}{10.1016/j.chom.2014.11.003}.

\leavevmode\hypertarget{ref-Tropini2018}{}%
16. \textbf{Tropini C}, \textbf{Moss EL}, \textbf{Merrill BD},
\textbf{Ng KM}, \textbf{Higginbottom SK}, \textbf{Casavant EP},
\textbf{Gonzalez CG}, \textbf{Fremin B}, \textbf{Bouley DM},
\textbf{Elias JE}, \textbf{Bhatt AS}, \textbf{Huang KC},
\textbf{Sonnenburg JL}. 2018. Transient osmotic perturbation causes
long-term alteration to the gut microbiota. Cell
\textbf{173}:1742--1754.e17.
doi:\href{https://doi.org/10.1016/j.cell.2018.05.008}{10.1016/j.cell.2018.05.008}.

\leavevmode\hypertarget{ref-VanInsberghe2020}{}%
17. \textbf{VanInsberghe D}, \textbf{Elsherbini JA}, \textbf{Varian B},
\textbf{Poutahidis T}, \textbf{Erdman S}, \textbf{Polz MF}. 2020.
Diarrhoeal events can trigger long-term clostridium difficile
colonization with recurrent blooms. Nature Microbiology
\textbf{5}:642--650.
doi:\href{https://doi.org/10.1038/s41564-020-0668-2}{10.1038/s41564-020-0668-2}.

\leavevmode\hypertarget{ref-Olson2013}{}%
18. \textbf{Olson A}, \textbf{Diebel LN}, \textbf{Liberati DM}. 2013.
Effect of host defenses on clostridium difficile toxininduced intestinal
barrier injury. Journal of Trauma and Acute Care Surgery
\textbf{74}:983--990.
doi:\href{https://doi.org/10.1097/ta.0b013e3182858477}{10.1097/ta.0b013e3182858477}.

\leavevmode\hypertarget{ref-Diebel2014}{}%
19. \textbf{Diebel LN}, \textbf{Liberati DM}. 2014. Reinforcement of the
intestinal mucus layer protects against clostridium difficile intestinal
injury in~vitro. Journal of the American College of Surgeons
\textbf{219}:460--468.
doi:\href{https://doi.org/10.1016/j.jamcollsurg.2014.05.005}{10.1016/j.jamcollsurg.2014.05.005}.

\leavevmode\hypertarget{ref-vanNood2013}{}%
20. \textbf{Nood E van}, \textbf{Vrieze A}, \textbf{Nieuwdorp M},
\textbf{Fuentes S}, \textbf{Zoetendal EG}, \textbf{Vos WM de},
\textbf{Visser CE}, \textbf{Kuijper EJ}, \textbf{Bartelsman JFWM},
\textbf{Tijssen JGP}, \textbf{Speelman P}, \textbf{Dijkgraaf MGW},
\textbf{Keller JJ}. 2013. Duodenal infusion of donor feces for
RecurrentClostridium difficile. New England Journal of Medicine
\textbf{368}:407--415.
doi:\href{https://doi.org/10.1056/nejmoa1205037}{10.1056/nejmoa1205037}.

\leavevmode\hypertarget{ref-Razik2017}{}%
21. \textbf{Razik R}, \textbf{Osman M}, \textbf{Lieberman A},
\textbf{Allegretti JR}, \textbf{Kassam Z}. 2017. Faecal microbiota
transplantation for clostridium difficile infection: A multicentre study
of non-responders. Medical Journal of Australia \textbf{207}:159--160.
doi:\href{https://doi.org/10.5694/mja16.01452}{10.5694/mja16.01452}.

\leavevmode\hypertarget{ref-Postigo2012}{}%
22. \textbf{Postigo R}, \textbf{Kim JH}. 2012. Colonoscopic versus
nasogastric fecal transplantation for the treatment of clostridium
difficile infection: A review and pooled analysis. Infection
\textbf{40}:643--648.
doi:\href{https://doi.org/10.1007/s15010-012-0307-9}{10.1007/s15010-012-0307-9}.

\leavevmode\hypertarget{ref-Liacouras1996}{}%
23. \textbf{Liacouras CA}, \textbf{Piccoli DA}. 1996. Whole-bowel
irrigation as an adjunct to the treatment of chronic, relapsing
clostridium difficile colitis. Journal of Clinical Gastroenterology
\textbf{22}:186--189.
doi:\href{https://doi.org/10.1097/00004836-199604000-00007}{10.1097/00004836-199604000-00007}.

\leavevmode\hypertarget{ref-Schubert2015}{}%
24. \textbf{Schubert AM}, \textbf{Sinani H}, \textbf{Schloss PD}. 2015.
Antibiotic-induced alterations of the murine gut microbiota and
subsequent effects on colonization resistance against \emph{Clostridium
difficile}. mBio \textbf{6}.
doi:\href{https://doi.org/10.1128/mbio.00974-15}{10.1128/mbio.00974-15}.

\leavevmode\hypertarget{ref-Jenior2017}{}%
25. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2017. \emph{Clostridium difficile} colonizes
alternative nutrient niches during infection across distinct murine gut
microbiomes. mSystems \textbf{2}.
doi:\href{https://doi.org/10.1128/msystems.00063-17}{10.1128/msystems.00063-17}.

\leavevmode\hypertarget{ref-Tomkovich2020}{}%
26. \textbf{Tomkovich S}, \textbf{Stough JMA}, \textbf{Bishop L},
\textbf{Schloss PD}. 2020. The initial gut microbiota and response to
antibiotic perturbation influence clostridioides difficile clearance in
mice. mSphere \textbf{5}.
doi:\href{https://doi.org/10.1128/msphere.00869-20}{10.1128/msphere.00869-20}.

\leavevmode\hypertarget{ref-Nguyen2015}{}%
27. \textbf{Nguyen TLA}, \textbf{Vieira-Silva S}, \textbf{Liston A},
\textbf{Raes J}. 2015. How informative is the mouse for human gut
microbiota research? Disease Models \& Mechanisms \textbf{8}:1--16.
doi:\href{https://doi.org/10.1242/dmm.017400}{10.1242/dmm.017400}.

\leavevmode\hypertarget{ref-Reeves2011}{}%
28. \textbf{Reeves AE}, \textbf{Theriot CM}, \textbf{Bergin IL},
\textbf{Huffnagle GB}, \textbf{Schloss PD}, \textbf{Young VB}. 2011. The
interplay between microbiome dynamics and pathogen dynamics in a murine
model of \emph{Clostridium difficile} infection \textbf{2}:145--158.
doi:\href{https://doi.org/10.4161/gmic.2.3.16333}{10.4161/gmic.2.3.16333}.

\leavevmode\hypertarget{ref-Dieterle2020}{}%
29. \textbf{Dieterle MG}, \textbf{Putler R}, \textbf{Perry DA},
\textbf{Menon A}, \textbf{Abernathy-Close L}, \textbf{Perlman NS},
\textbf{Penkevich A}, \textbf{Standke A}, \textbf{Keidan M},
\textbf{Vendrov KC}, \textbf{Bergin IL}, \textbf{Young VB}, \textbf{Rao
K}. 2020. Systemic inflammatory mediators are effective biomarkers for
predicting adverse outcomes in clostridioides difficile infection. mBio
\textbf{11}.
doi:\href{https://doi.org/10.1128/mbio.00180-20}{10.1128/mbio.00180-20}.

\leavevmode\hypertarget{ref-Jenior2018}{}%
30. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2018. \emph{Clostridium difficile} alters the
structure and metabolism of distinct cecal microbiomes during initial
infection to promote sustained colonization. mSphere \textbf{3}.
doi:\href{https://doi.org/10.1128/msphere.00261-18}{10.1128/msphere.00261-18}.

\leavevmode\hypertarget{ref-Vornhagen2020}{}%
31. \textbf{Vornhagen J}, \textbf{Bassis CM}, \textbf{Ramakrishnan S},
\textbf{Hein R}, \textbf{Mason S}, \textbf{Bergman Y}, \textbf{Sunshine
N}, \textbf{Fan Y}, \textbf{Timp W}, \textbf{Schatz MC}, \textbf{Young
VB}, \textbf{Simner PJ}, \textbf{Bachman MA}. 2020. A plasmid locus
associated with klebsiella clinical infections encodes a
microbiome-dependent gut fitness factor.
doi:\href{https://doi.org/10.1101/2020.02.26.963322}{10.1101/2020.02.26.963322}.

\leavevmode\hypertarget{ref-Dieterle2018}{}%
32. \textbf{Dieterle MG}, \textbf{Rao K}, \textbf{Young VB}. 2018. Novel
therapies and preventative strategies for primary and
recurrentClostridium difficileinfections. Annals of the New York Academy
of Sciences \textbf{1435}:110--138.
doi:\href{https://doi.org/10.1111/nyas.13958}{10.1111/nyas.13958}.

\leavevmode\hypertarget{ref-Theriot2011}{}%
33. \textbf{Theriot CM}, \textbf{Koumpouras CC}, \textbf{Carlson PE},
\textbf{Bergin II}, \textbf{Aronoff DM}, \textbf{Young VB}. 2011.
Cefoperazone-treated mice as an experimental platform to assess
differential virulence ofClostridium difficilestrains. Gut Microbes
\textbf{2}:326--334.
doi:\href{https://doi.org/10.4161/gmic.19142}{10.4161/gmic.19142}.

\end{document}